---
title: "Simulation Study for Timing Model"
author: "James Watson"
output:
  html_document:
    df_print: paged
    keep_md: yes
    toc: yes
---

# Preamble 

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE, cache.comments = FALSE, 
                      echo = F, include = TRUE, 
                      fig.width = 7, fig.height = 7,
                      fig.pos = 'H', 
                      dev = 'png', dpi = 300)

load('../RData/RPackages_List.RData')
new.pkg <- pkgs[!(pkgs %in% installed.packages()[, "Package"])]
if(length(new.pkg) > 0){ # Install using .R script
  stop('Please run the script Install_and_load_required_packages.R before returning to this script. Thanks.')
} else{
  sapply(pkgs, require, character.only = TRUE) # Load all packages
}

source('Functions.R')
```


```{r, echo=FALSE}
# Note that recompilation is necessary on different platforms
RECOMPILE_MODELS = F
# Running the full stan models can take several days
RUN_MODELS = F
```

Global options for stan model runs
```{r, echo=TRUE}
Chains = 4
options(mc.cores = Chains)
IT = 10^4
WarmUp = .5*IT
thin = 40
Sim_Iterations = 50
```

Global simulation sample sizes and follow-up time
```{r, echo=TRUE}
N_PMQ = 800
N_AS = 200
N_CQ = 200
FUP_time = 300
```

# Data simulation functions

We explore a few scenarios of increasing complexity to check whether:

* The model can recover correct parameters when the data generating process is correctly specified (sanity check)
* The effect of a mis-specified data generating process. For this we look at the impact of seasonality


## Data simulation for model 1

This simulates data under the assumptions of Model 1.

We set the simulation parameters:
```{r, echo=TRUE}
params_M1 = list(lambda = 1/1200,
                 gamma = 1/80,
                 lambda_recrud = 1/10,
                 EarlyL = 0.5,
                 c1_CQ = 0.01,
                 c1_AS = 0.01,
                 logit_mean_p = logit(0.2),
                 logit_sd_p = (-logit(0.2)+logit(0.8))/1.96,
                 AS_shape = 10,
                 AS_scale = 21,
                 CQ_shape = 10,
                 CQ_scale = 42,
                 rate_decrease = .66)
```


Generate data with these parameters:
```{r, echo=TRUE}
set.seed(475732)
out1 = simulate_dataset(N_PMQ = N_PMQ, N_CQ = N_CQ, 
                        N_AS = N_AS,FUP_time = FUP_time,
                        data_generation_function = generate_patient_data_Model1,
                        params = params_M1,
                        Study_Period = c(rep(1,N_PMQ/2),rep(2,N_PMQ/2 + N_AS + N_CQ)))
Simdata_Model1 = out1$Simdata
Simulation_truth1 = out1$Simulation_truth
```


# Run Stan Model 1 on simulated data generated from Model 1

Load or compile the stan model 1
```{r}
if(RECOMPILE_MODELS){
  writeLines('Compiling model 1....')
  source('../Timing_Model/StanModel1.R')
  save(Timing_Model1, file = '../RData/TimingModel/Timing_Model1.RData')
} else {
  load('../RData/TimingModel/Timing_Model1.RData')
}
```

Prior specification
```{r}
# The hierachical parameters defining the prior distributions for model 1
Prior_params_M1 = list(Hyper_lambda_shape = 100,
                       Hyper_lambda_rate = 100*(1/params_M1$lambda),
                       Hyper_gamma_shape = 100,
                       Hyper_gamma_rate = 100*(1/params_M1$gamma),
                       Hyper_lambda_recrud_shape = 100,
                       Hyper_lambda_recrud_rate = 100*(1/params_M1$lambda_recrud),
                       Hyper_AS_shape_mean = params_M1$AS_shape,
                       Hyper_AS_shape_sd = 1,
                       Hyper_AS_scale_mean = params_M1$AS_scale,
                       Hyper_AS_scale_sd = 1,
                       Hyper_CQ_shape_mean = params_M1$CQ_shape,
                       Hyper_CQ_shape_sd = 1,
                       Hyper_CQ_scale_mean = params_M1$CQ_scale,
                       Hyper_CQ_scale_sd = 2,
                       Hyper_logit_mean_p_mean = params_M1$logit_mean_p,
                       Hyper_logit_mean_p_sd = 1,
                       Hyper_logit_sd_p_lambda = 1,
                       Hyper_logit_c1_mean = logit(params_M1$c1_AS),
                       Hyper_logit_c1_sd = .25,
                       Early_L_logit_mean = logit(params_M1$EarlyL),
                       Early_L_logit_sd = .5,
                       Hyper_mean_rate_decrease = params_M1$rate_decrease,
                       Hyper_sd_rate_decrease = 0.25)
```



Fit stan model to simulated data
```{r, include=FALSE}
if(RUN_MODELS){
  tic()
  mod1_Fit = sampling(Timing_Model1,
                      data = c(Simdata_Model1, Prior_params_M1),
                      iter = IT, warmup = WarmUp,
                      chains = Chains, thin = thin, 
                      init = lapply(1:Chains, 
                                    FUN = function(x,Prior_params_M1,N_noPMQ){
                                      generate_inital_values_M1(Prior_params_M1,N_noPMQ)
                                    }, 
                                    Prior_params_M1 = Prior_params_M1, 
                                    N_noPMQ = Simdata_Model1$N_noPMQ))
  save(mod1_Fit, file = '../RData/LargeFiles/StanModels_Sim1_mod1.RData')
  toc()
} else {
  load(file = '../RData/LargeFiles/StanModels_Sim1_mod1.RData')
}
thetas_mod1 = extract(mod1_Fit)
```

Plot output, comparing prior, ground truth and model estimate. This gives a qualitative assessement for a single simulation run.

```{r plotModel1_fit}
plot_output_model1(thetas = thetas_mod1,
                   Simulation_truth = Simulation_truth1,
                   Simdata_Model = Simdata_Model1,
                   Prior_params_M1,
                   params_M1 = params_M1)
```


#### Multiple random simulations to assess systematic bias 

We run 50 iterations of:

1. Generate a random dataset under the model assumptions
2. Fit the model to the data

This allows us to compare model estimates versus true values to detect systematic bias.


```{r, include=FALSE}
params_interest = c('lambda','gamma','logit_mean_p','rate_decrease',
                    'logit_sd_p','logit_EarlyL')
if(RUN_MODELS){
  
  comparison_matrix = array(dim = c(Sim_Iterations,length(params_interest),2))
  for(sim in 1:Sim_Iterations){
    
    results = full_shebang(simulation_patient_function = generate_patient_data_Model1,
                           simulate_data_function = simulate_dataset,
                           stan_model = Timing_Model1,
                           params = params_M1,
                           init_values_function = generate_inital_values_M1,
                           N_AS = N_AS, N_CQ = N_CQ, N_PMQ = N_PMQ,
                           Prior_params = Prior_params_M1,
                           IT = IT,WarmUp = WarmUp,Chains = 1,thin = thin,
                           params_interest = params_interest,
                           Study_Period = c(rep(1,N_PMQ/2),rep(2,N_PMQ/2 + N_AS + N_CQ)))
    
    # record results: mean estimated value and true simulation value
    for(p in params_interest){
      comparison_matrix[sim,which(p==params_interest),] = results$summary_parameters[[p]]
    }
  }
  
  save(comparison_matrix,file = '../RData/TimingModel/simulation_full_results_mod1.RData')
} else {
  load('../RData/TimingModel/simulation_full_results_mod1.RData')
}
```

```{r full_shebang_results_mod1}
par(las=1,mfrow=c(3,2))
for(p in params_interest){
  hist(comparison_matrix[,which(p==params_interest),1], main=p)
  if(p=='logit_EarlyL'){
    abline(v= logit(params_M1$EarlyL),lwd=3,col='red')
  } else {
    abline(v=unique(comparison_matrix[,which(p==params_interest),2]),
           lwd=3,col='red')
  }
}
```

## Data simulation for model 2

Model 2 does not assume 100% efficacy of primaquine.

Ground truth model parameters for simulated data:
```{r}
params_M2 = c(params_M1,
              logit_mean_p_PMQ = logit(0.95),
              logit_sd_p_PMQ = (-logit(0.8)+logit(0.99))/1.96)
```

Generate simulated data under the assumptions of model 2
```{r}
set.seed(7656728)
out2 = simulate_dataset(N_PMQ = N_PMQ, 
                        N_CQ = N_CQ, 
                        N_AS = N_AS,
                        FUP_time = FUP_time,
                        Study_Period = c(rep(1,N_PMQ/2),rep(2,N_PMQ/2 + N_AS + N_CQ)),
                        data_generation_function = generate_patient_data_Model2,
                        params = params_M2)
Simdata_Model2 = out2$Simdata
Simulation_truth2 = out2$Simulation_truth
table(Simulation_truth2$True_state, Simulation_truth2$Drug)
```

Prior specification for stan fit:
```{r}
Prior_params_M2 = c(Prior_params_M1,
                    Hyper_logit_mean_pPMQ_mean = logit(0.95),
                    Hyper_logit_mean_pPMQ_sd = (logit(0.95)-logit(0.7))/1.96,
                    Hyper_logit_sd_pPMQ_lambda = 1)
```


Load or compile the stan model 2
```{r}
if(RECOMPILE_MODELS){
  writeLines('Compiling model 2....')
  source('../Timing_Model/StanModel2.R')
  save(Timing_Model2, file = '../RData/TimingModel/Timing_Model2.RData')
} else {
  load('../RData/TimingModel/Timing_Model2.RData')
}
```


Fit stan model 2 to simulated data from model 2
```{r, include=FALSE}
if(RUN_MODELS){
  tic()
  mod2_Fit = sampling(Timing_Model2,
                      data = c(Simdata_Model2, Prior_params_M2),
                      iter = IT, warmup = WarmUp,
                      chains=Chains, thin = thin, 
                      init = lapply(1:Chains, 
                                    FUN = function(x,Prior_params_M2,N_noPMQ,N_PMQ){
                                      generate_inital_values_M2(Prior_params_M2,N_noPMQ,N_PMQ)
                                    }, 
                                    Prior_params_M2 = Prior_params_M2, 
                                    N_noPMQ = Simdata_Model2$N_noPMQ,
                                    N_PMQ = Simdata_Model2$N_PMQ))
  save(mod2_Fit, file = '../RData/LargeFiles/StanModels_Sim2_mod2.RData')
  toc()
} else {
  load(file = '../RData/LargeFiles/StanModels_Sim2_mod2.RData')
}
thetas_mod2 = extract(mod2_Fit)
```


Plot output, comparing prior, ground truth and model estimate:
```{r plotModel2_fit}
plot_output_model2(thetas = thetas_mod2,Simulation_truth = Simulation_truth2,
                   Simdata_Model = Simdata_Model2,Prior_params_M2 = Prior_params_M2,
                   params_M2 = params_M2)
```

Compare summary statistics from simulated data and estimated:
```{r}
classification_mod2_upper = apply(thetas_mod2$prob_labels,c(2,3),quantile, probs=0.975)
classification_mod2 = apply(thetas_mod2$prob_labels,c(2,3),median)
classification_mod2_lower = apply(thetas_mod2$prob_labels,c(2,3),quantile, probs=0.025)

writeLines(sprintf('The true failure rate in the simulated data is %s%%. Model 2 estimates the failure rate to be %s%% (95%% CI: %s-%s)',
                   # the true failure rate in the primaquine arm
                   round(100*sum(Simulation_truth2$True_state[Simulation_truth2$Drug=="CHQ/PMQ"] != 'Reinfection')/sum(Simulation_truth2$Drug=="CHQ/PMQ"),1),
                   round(100*(1-sum(classification_mod2[Simdata_Model2$Drug==2,1])/sum(Simdata_Model2$Drug==2)),1),
                   round(100*(1-sum(classification_mod2_upper[Simdata_Model2$Drug==2,1])/sum(Simdata_Model2$Drug==2)),1),
                   round(100*(1-sum(classification_mod2_lower[Simdata_Model2$Drug==2,1])/sum(Simdata_Model2$Drug==2)),1)))
```


#### Multiple random simulations to assess systematic bias 

```{r, include=FALSE}
params_interest = c('lambda','gamma','logit_mean_p','logit_mean_p_PMQ',
                    'logit_sd_p','logit_EarlyL')
if(RUN_MODELS){
  
  comparison_matrix = array(dim = c(Sim_Iterations,length(params_interest),2))
  failure_matrix = array(dim = c(Sim_Iterations, 2))
  for(sim in 1:Sim_Iterations){
    
    results = full_shebang(simulation_patient_function = generate_patient_data_Model2,
                           simulate_data_function = simulate_dataset,
                           stan_model = Timing_Model2,
                           params = params_M2,
                           init_values_function = generate_inital_values_M2,
                           N_AS = N_AS, N_CQ = N_CQ, N_PMQ = N_PMQ,
                           Prior_params = Prior_params_M2,
                           IT = IT,WarmUp = WarmUp,Chains = Chains,thin = thin,
                           params_interest = params_interest,
                           Study_Period = c(rep(1,N_PMQ/2),rep(2,N_PMQ/2 + N_AS + N_CQ)))
    
    # record results: mean estimated value and true simulation value
    for(p in params_interest){
      comparison_matrix[sim,which(p==params_interest),] = results$summary_parameters[[p]]
    }
    failure_matrix[sim,] = results$summary_stats
  }
  simulation_full_results_mod2 = list(comparison_matrix=comparison_matrix,
                                      failure_matrix=failure_matrix)
  save(simulation_full_results_mod2,
       file = '../RData/TimingModel/simulation_full_results_mod2.RData')
} else {
  load('../RData/TimingModel/simulation_full_results_mod2.RData')
}
```

```{r full_shebang_results_mod2}
par(las=1, mfrow=c(3,2))
for(p in params_interest){
  hist(simulation_full_results_mod2$comparison_matrix[,which(p==params_interest),1], main=p)
  abline(v=unique(simulation_full_results_mod2$comparison_matrix[,which(p==params_interest),2]),col='red',lwd=3)
}
par(mfrow=c(1,1))
plot(100*simulation_full_results_mod2$failure_matrix[,1],
     100*simulation_full_results_mod2$failure_matrix[,2],
     xlab='True PMQ failure rate in simulated data', ylab = 'Estimated PMQ failure rate')
lines(c(0,100),c(0,100))
```


# Simulating seasonal variation

Estimate the empirical seasonal reinfecton distribution from enrollment episodes. This assumes that the majority of enrollment episodes are reinfections which should be approximately true.
```{r}
load('../RData/TimingModel/Combined_Time_Event.RData')
Combined_Time_Data = filter(Combined_Time_Data, Censored == -1)
ind = Combined_Time_Data$WeekTime>52
Combined_Time_Data$WeekTime[ind] = Combined_Time_Data$WeekTime[ind]-52
hist(Combined_Time_Data$WeekTime, breaks = 0:52, 
     main = 'Week of enrollment',xlab='Week of the year')
# the vector we can use for empirical distribution sampling
seasonal_sampling_vector = Combined_Time_Data$WeekTime
```

Generate patient data under the assumptions of model 1 with the additional assumption of seasonality for reinfection.
```{r}
# test to show this does the right thing
set.seed(7576)
ys = sapply(1:1000, FUN = function(x, seasonal_sampling_vector){
  generate_reinfection_time_seasonal(params = params_M1,
                                     week_time = 0,
                                     seasonal_sampling_vector = seasonal_sampling_vector,
                                     Study_Period = 1)
}, seasonal_sampling_vector)
hist(ys, breaks = seq(0,20000,by=14), xlim = c(0,3*360), 
     main='Reinfection times',
     xlab='days from 1st January')
```


Generate data with these parameters:
```{r}
set.seed(87678)
tic()
out3 = simulate_dataset(N_PMQ = N_PMQ, N_CQ = N_CQ, 
                        N_AS = N_AS,FUP_time = FUP_time,
                        data_generation_function = generate_patient_data_Model2_Seasonal,
                        params = params_M2,
                        seasonal_sampling_vector = seasonal_sampling_vector,
                        Study_Period = c(rep(1,N_PMQ/2),rep(2,N_PMQ/2 + N_AS + N_CQ)))
toc()
Simdata_Model_Seasonal = out3$Simdata
Simulation_truth_Seasonal = out3$Simulation_truth
```


Fit stan model to simulated data where assumption of constant reinfection rate is wrong
```{r, include=FALSE}
if(RUN_MODELS){
  mod_seasonal_Fit = sampling(Timing_Model2,
                              data = c(Simdata_Model_Seasonal, Prior_params_M2),
                              iter = IT, warmup = WarmUp,
                              chains=Chains, thin = thin, 
                              init = lapply(1:Chains, 
                                            FUN = function(x,Prior_params_M2,N_noPMQ,N_PMQ){
                                              generate_inital_values_M2(Prior_params_M2,N_noPMQ,N_PMQ)
                                            }, 
                                            Prior_params = Prior_params_M2, 
                                            N_noPMQ = Simdata_Model_Seasonal$N_noPMQ,
                                            N_PMQ = Simdata_Model_Seasonal$N_PMQ))
  save(mod_seasonal_Fit, file = '../RData/LargeFiles/StanModels_SimSeasonal_mod2.RData')
} else {
  load(file = '../RData/LargeFiles/StanModels_SimSeasonal_mod2.RData')
}
thetas_mod_seasonal = extract(mod_seasonal_Fit)
```


Plot output, comparing prior, ground truth and model estimate:
```{r plotModel_Seasonal_fit}
plot_output_model2(thetas = thetas_mod_seasonal,
                   Simulation_truth = Simulation_truth_Seasonal,
                   Simdata_Model = Simdata_Model_Seasonal,
                   Prior_params_M2 = Prior_params_M2, 
                   params_M2 = params_M2)
```

### Run multiple random iterations to assess model fit and bias


```{r, include=FALSE}
params_interest = c('lambda','gamma','logit_mean_p','rate_decrease',
                    'logit_sd_p','logit_EarlyL')
if(RUN_MODELS){
  comparison_matrix = array(dim = c(Sim_Iterations,length(params_interest),2))
  failure_matrix = array(dim = c(Sim_Iterations, 2))
  for(sim in 1:Sim_Iterations){
    results = full_shebang(simulation_patient_function = 
                             generate_patient_data_Model2_Seasonal,
                           simulate_data_function = simulate_dataset,
                           stan_model = Timing_Model2,
                           params = params_M2,
                           init_values_function = generate_inital_values_M2,
                           N_AS = N_AS, N_CQ = N_CQ, N_PMQ = N_PMQ,
                           Prior_params = Prior_params_M2,
                           IT = IT, WarmUp = WarmUp, Chains = Chains, thin = thin,
                           params_interest = params_interest,
                           seasonal_sampling_vector=seasonal_sampling_vector,
                           Study_Period = c(rep(1,N_PMQ/2),rep(2,N_PMQ/2 + N_AS + N_CQ)))
    
    # record results: mean estimated value and true simulation value
    for(p in params_interest){
      comparison_matrix[sim,which(p==params_interest),] = results$summary_parameters[[p]]
    }
    failure_matrix[sim,] = results$summary_stats
  }
  res = list(comparison_matrix=comparison_matrix, failure_matrix=failure_matrix)
  save(res,
       file = '../RData/TimingModel/simulation_full_results_mod_seasonal.RData')
} else {
  load('../RData/TimingModel/simulation_full_results_mod_seasonal.RData')
}
```



```{r full_shebang_results_mod_seasonal}
par(las=1, mfrow=c(2,3))
for(p in params_interest){
  
  hist(res$comparison_matrix[,which(p==params_interest),1], main=p)
  if(p == 'logit_EarlyL'){
    abline(v=logit(params_M2$EarlyL), col='red',lwd=3)
    writeLines(sprintf('For parameter %s, %s%% are below the true parameter',
                       p,100*sum(res$comparison_matrix[,which(p==params_interest),1]<                                                                                  logit(params_M2$EarlyL))/Sim_Iterations))
  } else {
    abline(v=unique(res$comparison_matrix[,which(p==params_interest),2]),col='red',lwd=3)
    writeLines(sprintf('For parameter %s, %s%% are below the true parameter',
                       p,100*sum(res$comparison_matrix[,which(p==params_interest),1]<                                                                                  unique(res$comparison_matrix[,which(p==params_interest),2]))/Sim_Iterations))
  }
}
par(mfrow=c(1,1))
plot(100*res$failure_matrix[,1],100*res$failure_matrix[,2],
     xlab='True PMQ failure rate',ylab = 'Estimated PMQ failure rate')
lines(c(0,100),c(0,100))
writeLines(sprintf('%s%% of the simulations overestimate the failure rate of PMQ',
                   100*sum(res$failure_matrix[,1]< res$failure_matrix[,2])/nrow(res$failure_matrix)))
```


# Posterior predictive p-values

In this section we simulate data from the posterior predictive for Model 2. We then apply a qualitative model check by comparing summary statistics from the posterior predicitive with the true data summary statistics.

```{r}
load('../RData/TimingModel/Posterior_theta_samples.RData')
load('../RData/TimingModel/Combined_Time_Event.RData')

PMQ_FUP = Combined_Time_Data$FU_time[!duplicated(Combined_Time_Data$patientid)&
                                       Combined_Time_Data$arm_num=='CHQ/PMQ']
CQ_FUP = Combined_Time_Data$FU_time[!duplicated(Combined_Time_Data$patientid)&
                                      Combined_Time_Data$arm_num=='CHQ']
AS_FUP = Combined_Time_Data$FU_time[!duplicated(Combined_Time_Data$patientid)&
                                      Combined_Time_Data$arm_num=='AS']
N_PMQ = sum(Combined_Time_Data$arm_num=='CHQ/PMQ'&
              !duplicated(Combined_Time_Data$patientid))
N_CQ = sum(Combined_Time_Data$arm_num=='CHQ'&
             !duplicated(Combined_Time_Data$patientid))
N_AS = sum(Combined_Time_Data$arm_num=='AS'&
             !duplicated(Combined_Time_Data$patientid))
N_PMQ_BPD = sum(Combined_Time_Data$arm_num=='CHQ/PMQ'&
                  Combined_Time_Data$Study_Period==2 &
                  !duplicated(Combined_Time_Data$patientid))

set.seed(4993758)
params_M2_posterior = list()
i = sample(1:nrow(thetas2_matrix),1)
colnames(thetas2_matrix)[colnames(thetas2_matrix)=='logit_EarlyL']='EarlyL'
colnames(thetas2_matrix)[colnames(thetas2_matrix)=='logit_c1_AS']='c1_AS'
colnames(thetas2_matrix)[colnames(thetas2_matrix)=='logit_c1_CQ']='c1_CQ'
for(p in names(params_M2)){
  params_M2_posterior[[p]] = thetas2_matrix[i,p]
}
params_M2_posterior$EarlyL = inv.logit(params_M2_posterior$EarlyL)
params_M2_posterior$c1_CQ = inv.logit(params_M2_posterior$c1_CQ)
params_M2_posterior$c1_AS = inv.logit(params_M2_posterior$c1_AS)

dat_posterior =
  simulate_dataset(N_PMQ = N_PMQ,N_CQ = N_CQ,N_AS = N_AS,
                   FUP_time = c(sample(x = PMQ_FUP,size = length(PMQ_FUP),replace = T),
                                sample(x = CQ_FUP,size = length(CQ_FUP),replace = T),
                                sample(x = AS_FUP,size = length(AS_FUP),replace = T)),
                   Study_Period = c(rep(2,N_PMQ_BPD),rep(2,N_PMQ-N_PMQ_BPD),rep(1,N_CQ+N_AS)),
                   data_generation_function = generate_patient_data_Model2,
                   params = params_M2_posterior)

AS_ind = dat_posterior$Simdata$Drug==0 & 
  dat_posterior$Simdata$Censored==0
AS_ind_true = Combined_Time_Data$arm_num=='AS'&
  Combined_Time_Data$Censored==0
CQ_ind = dat_posterior$Simdata$Drug==1 & 
  dat_posterior$Simdata$Censored==0
CQ_ind_true = Combined_Time_Data$arm_num=='CHQ'&
  Combined_Time_Data$Censored==0
PMQ_ind = dat_posterior$Simdata$Drug==2 & dat_posterior$Simdata$Censored==0
PMQ_ind_true = Combined_Time_Data$arm_num=='CHQ/PMQ'&
  Combined_Time_Data$Censored==0

par(las=1, mfrow=c(1,3))
#** Artesunate monotherapy: posterior predictive versus true data **
hist(Combined_Time_Data$Time_to_event[AS_ind_true], 
     main = 'AS monotherapy', 
     xlab='Days from last episode', col=alpha('blue',alpha = .4), 
     breaks = seq(0,700,by=7), xlim = c(0,300))
hist(dat_posterior$Simdata$Durations[AS_ind],
     add=T, breaks = seq(0,700,by=7), col=alpha('red',alpha = .4))

#** Chloroquine monotherapy: posterior predictive versus true data **
hist(Combined_Time_Data$Time_to_event[CQ_ind_true], main = 'CQ monotherapy', 
     xlab='days from last episode', col=alpha('blue',alpha = .4), 
     breaks = seq(0,700,by=7), xlim = c(0,300))
hist(dat_posterior$Simdata$Durations[CQ_ind],
     add=T, breaks = seq(0,700,by=7), col=alpha('red',alpha = .4))

#** Primaquine+ : posterior predictive versus true data **
hist(Combined_Time_Data$Time_to_event[PMQ_ind_true], main = 'PMQ+', 
     xlab='days from last episode', col=alpha('blue',alpha = .4), 
     breaks = seq(0,700,by=7), xlim = c(0,300))
hist(dat_posterior$Simdata$Durations[PMQ_ind],
     add=T, breaks = seq(0,700,by=7), col=alpha('red',alpha = .4))



# *************** Empirical cdfs **************
par(las=1, mfrow=c(1,3))
plot.ecdf(Combined_Time_Data$Time_to_event[AS_ind_true],
          col='blue',verticals = T, xlab='Days from last episode',
          main='Recurrence after AS monotherapy')
plot.ecdf(dat_posterior$Simdata$Durations[AS_ind],
          add=T,col='red',verticals = T)

plot.ecdf(Combined_Time_Data$Time_to_event[CQ_ind_true],
          col='blue',verticals = T, xlab='Days from last episode',
          main='Recurrence after CQ monotherapy')
plot.ecdf(dat_posterior$Simdata$Durations[CQ_ind],
          add=T,col='red',verticals = T)


plot.ecdf(Combined_Time_Data$Time_to_event[PMQ_ind_true],
          col='blue',verticals = T, xlab='Days from last episode',
          main='Recurrence after PMQ+')
plot.ecdf(dat_posterior$Simdata$Durations[PMQ_ind],
          add=T,col='red',verticals = T)
legend('bottomright',col=c('red','blue'),
       legend = c('Posterior predictive \n(simulated)',
                  'Observed trial data'),lwd=2, inset = 0.01)
```

```{r }
Sim_Iterations = 500
if(RUN_MODELS){
  
  res = array(dim = c(Sim_Iterations,3))
  pb = txtProgressBar(min=1, max = Sim_Iterations,style = 3)
  for(Sim in 1:Sim_Iterations){
    
    params_M2_posterior = list()
    i = sample(1:nrow(thetas2_matrix),1)
    for(p in names(params_M2)){
      params_M2_posterior[[p]] = thetas2_matrix[i,p]
    }
    params_M2_posterior$EarlyL = inv.logit(params_M2_posterior$EarlyL)
    params_M2_posterior$c1_CQ = inv.logit(params_M2_posterior$c1_CQ)
    params_M2_posterior$c1_AS = inv.logit(params_M2_posterior$c1_AS)
    
    dat_posterior =
      simulate_dataset(N_PMQ = N_PMQ,N_CQ = N_CQ,N_AS = N_AS,
                       FUP_time = c(sample(x = PMQ_FUP,size = length(PMQ_FUP),replace = T),
                                    sample(x = CQ_FUP,size = length(CQ_FUP),replace = T),
                                    sample(x = AS_FUP,size = length(AS_FUP),replace = T)),
                       Study_Period = c(rep(2,N_PMQ_BPD),rep(2,N_PMQ-N_PMQ_BPD),rep(1,N_CQ+N_AS)),
                       data_generation_function = generate_patient_data_Model2,
                       params = params_M2_posterior)
    
    AS_ind = dat_posterior$Simdata$Drug==0 & 
      dat_posterior$Simdata$Censored==0
    CQ_ind = dat_posterior$Simdata$Drug==1 & 
      dat_posterior$Simdata$Censored==0
    PMQ_ind = dat_posterior$Simdata$Drug==2 & dat_posterior$Simdata$Censored==0
    
    # compute recurrences per follow-up year
    res[Sim,1] = length(dat_posterior$Simdata$Durations[AS_ind])/
      (sum(dat_posterior$Simdata$Durations[dat_posterior$Simdata$Drug==0])/365)
    res[Sim,2] = length(dat_posterior$Simdata$Durations[CQ_ind])/
      (sum(dat_posterior$Simdata$Durations[dat_posterior$Simdata$Drug==1])/365)
    res[Sim,3] = length(dat_posterior$Simdata$Durations[PMQ_ind])/
      (sum(dat_posterior$Simdata$Durations[dat_posterior$Simdata$Drug==2])/365)
    setTxtProgressBar(pb,value = Sim)
  }
  save(res, file = '../RData/TimingModel/PosteriorPredictivePvals.RData')
} else {
  load('../RData/TimingModel/PosteriorPredictivePvals.RData')
}
```

```{r posterior_pred_pvalues}
par(mfrow=c(1,3))
hist(res[,1], xlab = 'Recurrences per person-year (%)', main = 'Artesunate monotherapy')
true_stat_AS = length(Combined_Time_Data$Time_to_event[AS_ind_true])/
  (sum(Combined_Time_Data$Time_to_event[Combined_Time_Data$arm_num=='AS'])/365)
abline(v = true_stat_AS, col='blue',lwd=3)

hist(res[,2], xlab = 'Recurrences per person-year (%)', main = 'Chloroquine monotherapy')
true_stat_CQ = length(Combined_Time_Data$Time_to_event[CQ_ind_true])/
  (sum(Combined_Time_Data$Time_to_event[Combined_Time_Data$arm_num=='CHQ'])/365)
abline(v = true_stat_CQ,  col='blue',lwd=3)

hist(res[,3], xlab = 'Recurrences per person-year (%)', main = 'Primaquine+')
true_stat_PMQ = length(Combined_Time_Data$Time_to_event[PMQ_ind_true])/
  (sum(Combined_Time_Data$Time_to_event[Combined_Time_Data$arm_num=='CHQ/PMQ'])/365)
abline(v = true_stat_PMQ,col='blue',lwd=3)

PP_pval_AS = min(sum(res[,1]>true_stat_AS)/Sim_Iterations, 1-sum(res[,1]>true_stat_AS)/Sim_Iterations)*2
writeLines(sprintf('The posterior predictive p value for the number of recurrences per person-year in the AS arm is %s',PP_pval_AS))
PP_pval_CQ = min(sum(res[,2]>true_stat_CQ)/Sim_Iterations,1-sum(res[,2]>true_stat_CQ)/Sim_Iterations)*2
writeLines(sprintf('The posterior predictive p value for the number of recurrences per person-year in the CQ arm is %s',PP_pval_CQ))
PP_pval_PMQ = min(sum(res[,3]>true_stat_PMQ)/Sim_Iterations,1-sum(res[,3]>true_stat_PMQ)/Sim_Iterations)*2
writeLines(sprintf('The posterior predictive p value for the number of recurrences per person-year in the PMQ+ arm is %s',PP_pval_PMQ))
```


## Summary plot

This is a plot to summarise the simulation based model checking (sup material).

```{r summary_timing_model_check}
par(las=1, bty='n', mar=c(4,5,6,3))
layout(mat = matrix(data = c(1,1,1,2,2,2,
                             3,3,4,4,5,5), nrow = 2, byrow = T))

# Model 2 on well-specified data
load('../RData/TimingModel/simulation_full_results_mod2.RData')
plot(100*simulation_full_results_mod2$failure_matrix[,1],
     100*simulation_full_results_mod2$failure_matrix[,2], pch=16,
     xlab='True PMQ failure rate in simulated data', ylab = 'Estimated PMQ failure rate')
title('Well-specified model')
lines(c(0,100),c(0,100))

# Seasonal data - model misspecification
load('../RData/TimingModel/simulation_full_results_mod_seasonal.RData')
plot(100*res$failure_matrix[,1],100*res$failure_matrix[,2],pch=16,
     xlab='True PMQ failure rate in simulated data',ylab = 'Estimated PMQ failure rate')
lines(c(0,100),c(0,100))
title('Mis-specified model')

# Posterior model checks
load('../RData/TimingModel/PosteriorPredictivePvals.RData')
hist(res[,1], xlab = 'Recurrences per person-year (%)', main = 'Artesunate monotherapy',col = 'grey')
true_stat_AS = length(Combined_Time_Data$Time_to_event[AS_ind_true])/
  (sum(Combined_Time_Data$Time_to_event[Combined_Time_Data$arm_num=='AS'])/365)
abline(v = true_stat_AS, col='blue',lwd=3)

hist(res[,2], xlab = 'Recurrences per person-year (%)', main = 'Chloroquine monotherapy',col='grey')
true_stat_CQ = length(Combined_Time_Data$Time_to_event[CQ_ind_true])/
  (sum(Combined_Time_Data$Time_to_event[Combined_Time_Data$arm_num=='CHQ'])/365)
abline(v = true_stat_CQ,  col='blue',lwd=3)

hist(res[,3], xlab = 'Recurrences per person-year (%)', main = 'Primaquine+',col='grey')
true_stat_PMQ = length(Combined_Time_Data$Time_to_event[PMQ_ind_true])/
  (sum(Combined_Time_Data$Time_to_event[Combined_Time_Data$arm_num=='CHQ/PMQ'])/365)
abline(v = true_stat_PMQ,col='blue',lwd=3)
```

