---
title: "Simulation Study for Timing Model"
author: "James Watson"
date: "07/05/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE, cache.comments = FALSE, 
                      echo = TRUE, include = TRUE, 
                      fig.width = 7, fig.height = 7,
                      fig.pos = 'H', 
                      dev = c('pdf','png'), dpi = 300)

load('../RData/RPackages_List.RData')
new.pkg <- pkgs[!(pkgs %in% installed.packages()[, "Package"])]
if(length(new.pkg) > 0){ # Install using .R script
  stop('Please run the script Install_and_load_required_packages.R before returning to this script. Thanks.')
} else{
  sapply(pkgs, require, character.only = TRUE) # Load all packages
}

source('Functions.R')
```


```{r, echo=FALSE}
# Note that recompilation is necessary on different platforms
RECOMPILE_MODELS = F
# Running the full stan models can take several days
RUN_MODELS = F
```

Global options for stan model runs
```{r}
Chains = 3
options(mc.cores = Chains)
IT = 10^4
WarmUp = .5*IT
thin = 40
```

Global simulation sample sizes and follow-up time
```{r}
N_PMQ = 800
N_AS = 220
N_CQ = 220
FUP_time = 300
```

## Data simulation functions

We explore a few scenarios of increasing complexity to check:

* The model can recover correct parameters when the data generating process is correctly specified (sanity check)
* The effect of a mis-specified data generating process. For this we look at the impact of seasonality


## Data simulation for model 1

This simulates data under the assumptions of Model 1.

We set the simulation parameters:
```{r}
params_M1 = list(lambda = 1/1200,
                 gamma = 1/80,
                 lambda_recrud = 1/10,
                 logit_EarlyL = 0.5,
                 logit_c1_CQ = 0.01,
                 logit_c1_AS = 0.01,
                 logit_mean_p = logit(0.2),
                 logit_sd_p = (-logit(0.2)+logit(0.8))/1.96,
                 AS_shape = 10,
                 AS_scale = 21,
                 CQ_shape = 10,
                 CQ_scale = 42,
                 rate_decrease = .66)
```


Generate data with these parameters:
```{r}
set.seed(475732)
out1 = simulate_dataset(N_PMQ = N_PMQ, N_CQ = N_CQ, 
                        N_AS = N_AS,FUP_time = FUP_time,
                        data_generation_function = generate_patient_data_Model1,
                        params = params_M1,
                        Study_Period = c(rep(1,N_PMQ/2),rep(2,N_PMQ/2 + N_AS + N_CQ)))
Simdata_Model1 = out1$Simdata
Simulation_truth1 = out1$Simulation_truth
```


# Run Stan Model 1 on simulated data generated from Model 1

Load or compile the stan model 1
```{r}
if(RECOMPILE_MODELS){
  writeLines('Compiling model 1....')
  source('../Timing_Model/StanModel1.R')
  save(Timing_Model1, file = '../RData/TimingModel/Timing_Model1.RData')
} else {
  load('../RData/TimingModel/Timing_Model1.RData')
}
```

Prior specification
```{r}
# The hierachical parameters defining the prior distributions for model 1
Prior_params_M1 = list(Hyper_lambda_shape = 100,
                       Hyper_lambda_rate = 100*(1/params_M1$lambda),
                       Hyper_gamma_shape = 100,
                       Hyper_gamma_rate = 100*(1/params_M1$gamma),
                       Hyper_lambda_recrud_shape = 100,
                       Hyper_lambda_recrud_rate = 100*(1/params_M1$lambda_recrud),
                       Hyper_AS_shape_mean = params_M1$AS_shape,
                       Hyper_AS_shape_sd = 1,
                       Hyper_AS_scale_mean = params_M1$AS_scale,
                       Hyper_AS_scale_sd = 1,
                       Hyper_CQ_shape_mean = params_M1$CQ_shape,
                       Hyper_CQ_shape_sd = 1,
                       Hyper_CQ_scale_mean = params_M1$CQ_scale,
                       Hyper_CQ_scale_sd = 2,
                       Hyper_logit_mean_p_mean = params_M1$logit_mean_p,
                       Hyper_logit_mean_p_sd = 1,
                       Hyper_logit_sd_p_lambda = 1,
                       Hyper_logit_c1_mean = params_M1$logit_c1_AS,
                       Hyper_logit_c1_sd = .25,
                       Early_L_logit_mean = params_M1$logit_EarlyL,
                       Early_L_logit_sd = .5,
                       Hyper_mean_rate_decrease = params_M1$rate_decrease,
                       Hyper_sd_rate_decrease = 0.25)
```



Fit stan model to simulated data
```{r, include=FALSE}
if(RUN_MODELS){
  tic()
  mod1_Fit = sampling(Timing_Model1,
                      data = c(Simdata_Model1, Prior_params_M1),
                      iter = IT, warmup = WarmUp,
                      chains = Chains, thin = thin, 
                      init = lapply(1:Chains, 
                                    FUN = function(x,Prior_params_M1,N_noPMQ){
                                      generate_inital_values_M1(Prior_params_M1,N_noPMQ)
                                    }, 
                                    Prior_params_M1 = Prior_params_M1, 
                                    N_noPMQ = Simdata_Model1$N_noPMQ))
  save(mod1_Fit, file = '../RData/LargeFiles/StanModels_Sim1_mod1.RData')
  toc()
} else {
  load(file = '../RData/LargeFiles/StanModels_Sim1_mod1.RData')
}
thetas_mod1 = extract(mod1_Fit)
```

Plot output, comparing prior, ground truth and model estimate. This gives a qualitative assessement but for a single simulation run.
```{r plotModel1_fit}
plot_output_model1(thetas_mod1,Simulation_truth1,
                   Simdata_Model1,Prior_params_M1)
```


#### Multiple random simulations to assess systematic bias 

```{r, include=FALSE}
params_interest = c('lambda','gamma','logit_mean_p',
                    'logit_sd_p','logit_EarlyL')
Sim_Iterations = 100
if(RUN_MODELS){
  
  comparison_matrix = array(dim = c(Sim_Iterations,length(params_interest),2))
  for(sim in 1:Sim_Iterations){
    
    results = full_shebang(simulation_patient_function = generate_patient_data_Model1,
                           simulate_data_function = simulate_dataset,
                           stan_model = Timing_Model1,
                           params = params_M1,
                           init_values_function = generate_inital_values_M1,
                           N_AS = N_AS, N_CQ = N_CQ, N_PMQ = N_PMQ,
                           Prior_params = Prior_params_M1,
                           IT = IT,WarmUp = WarmUp,Chains = 1,thin = thin,
                           params_interest = params_interest)
    
    # record results: mean estimated value and true simulation value
    for(p in params_interest){
      comparison_matrix[sim,which(p==params_interest),] = results$summary_parameters[[p]]
    }
  }
  
  save(comparison_matrix,file = '../RData/TimingModel/simulation_full_results_mod1.RData')
} else {
  load('../RData/TimingModel/simulation_full_results_mod1.RData')
}
```

```{r full_shebang_results_mod1}
for(p in params_interest){
  hist(comparison_matrix[,which(p==params_interest),1], main=p)
  abline(v=unique(comparison_matrix[,which(p==params_interest),2]),
         lwd=3,col='red')
}
```

## Data simulation for model 2

Model 2 does not assume 100% efficacy of primaquine.

Ground truth model parameters for simulated data:
```{r}
params_M2 = c(params_M1,
              logit_mean_p_PMQ = logit(0.95),
              logit_sd_p_PMQ = (-logit(0.8)+logit(0.99))/1.96)
```

Generate simulated data under the assumptions of model 2
```{r}
set.seed(7656728)
out2 = simulate_dataset(N_PMQ = N_PMQ, 
                        N_CQ = N_CQ, 
                        N_AS = N_AS,
                        FUP_time = FUP_time,
                        Study_Period = c(rep(1,N_PMQ/2),rep(2,N_PMQ/2 + N_AS + N_CQ)),
                        data_generation_function = generate_patient_data_Model2,
                        params = params_M2)
Simdata_Model2 = out2$Simdata
Simulation_truth2 = out2$Simulation_truth
table(Simulation_truth2$True_state, Simulation_truth2$Drug)
```

Prior specification for stan fit:
```{r}
Prior_params_M2 = c(Prior_params_M1,
                    Hyper_logit_mean_pPMQ_mean = logit(0.95),
                    Hyper_logit_mean_pPMQ_sd = (logit(0.95)-logit(0.7))/1.96,
                    Hyper_logit_sd_pPMQ_lambda = 1)
```


Load or compile the stan model 2
```{r}
if(RECOMPILE_MODELS){
  writeLines('Compiling model 2....')
  source('../Timing_Model/StanModel2.R')
  save(Timing_Model2, file = '../RData/TimingModel/Timing_Model2.RData')
} else {
  load('../RData/TimingModel/Timing_Model2.RData')
}
```


Fit stan model 2 to simulated data from model 2
```{r, include=FALSE}
if(RUN_MODELS){
  tic()
  mod2_Fit = sampling(Timing_Model2,
                      data = c(Simdata_Model2, Prior_params_M2),
                      iter = IT, warmup = WarmUp,
                      chains=Chains, thin = thin, 
                      init = lapply(1:Chains, 
                                    FUN = function(x,Prior_params_M2,N_noPMQ,N_PMQ){
                                      generate_inital_values_M2(Prior_params_M2,N_noPMQ,N_PMQ)
                                    }, 
                                    Prior_params_M2 = Prior_params_M2, 
                                    N_noPMQ = Simdata_Model2$N_noPMQ,
                                    N_PMQ = Simdata_Model2$N_PMQ))
  save(mod2_Fit, file = '../RData/LargeFiles/StanModels_Sim2_mod2.RData')
  toc()
} else {
  load(file = '../RData/LargeFiles/StanModels_Sim2_mod2.RData')
}
thetas_mod2 = extract(mod2_Fit)
```


Plot output, comparing prior, ground truth and model estimate:
```{r plotModel2_fit}
plot_output_model2(thetas_mod2,Simulation_truth2,Simdata_Model2,Prior_params_M2)
```

Compare summary statistics from simulated data and estimated:
```{r}
# the true failure rate in the primaquine arm
sum(Simulation_truth2$True_state[Simulation_truth2$Drug=="CHQ/PMQ"] != 'Reinfection')/sum(Simulation_truth2$Drug=="CHQ/PMQ")

classification_mod2 = apply(thetas_mod2$prob_labels,c(2,3),quantile, probs=0.975)
1-sum(classification_mod2[Simdata_Model2$Drug==2,1])/sum(Simdata_Model2$Drug==2)

classification_mod2 = apply(thetas_mod2$prob_labels,c(2,3),median)
1-sum(classification_mod2[Simdata_Model2$Drug==2,1])/sum(Simdata_Model2$Drug==2)

classification_mod2 = apply(thetas_mod2$prob_labels,c(2,3),quantile, probs=0.025)
1-sum(classification_mod2[Simdata_Model2$Drug==2,1])/sum(Simdata_Model2$Drug==2)
```


#### Multiple random simulations to assess systematic bias 

```{r, include=FALSE}
params_interest = c('lambda','gamma','logit_mean_p','logit_mean_p_PMQ',
                    'logit_sd_p','logit_EarlyL')
Sim_Iterations = 100
if(RUN_MODELS){
  
  comparison_matrix = array(dim = c(Sim_Iterations,length(params_interest),2))
  failure_matrix = array(dim = c(Sim_Iterations, 2))
  for(sim in 1:Sim_Iterations){
    
    results = full_shebang(simulation_patient_function = generate_patient_data_Model2,
                           simulate_data_function = simulate_dataset,
                           stan_model = Timing_Model2,
                           params = params_M2,
                           init_values_function = generate_inital_values_M2,
                           N_AS = N_AS, N_CQ = N_CQ, N_PMQ = N_PMQ,
                           Prior_params = Prior_params_M2,
                           IT = IT,WarmUp = WarmUp,Chains = Chains,thin = thin,
                           params_interest = params_interest)
    
    # record results: mean estimated value and true simulation value
    for(p in params_interest){
      comparison_matrix[sim,which(p==params_interest),] = results$summary_parameters[[p]]
    }
    failure_matrix[sim,] = results$summary_stats
  }
  simulation_full_results_mod2 = list(comparison_matrix=comparison_matrix,
                                      failure_matrix=failure_matrix)
  save(simulation_full_results_mod2,
       file = '../RData/TimingModel/simulation_full_results_mod2.RData')
} else {
  load('../RData/TimingModel/simulation_full_results_mod2.RData')
}
```

```{r full_shebang_results_mod2}
for(p in params_interest){
  hist(simulation_full_results_mod2$comparison_matrix[,which(p==params_interest),1], main=p)
  abline(v=unique(simulation_full_results_mod2$comparison_matrix[,which(p==params_interest),2]),col='red',lwd=3)
}

plot(100*simulation_full_results_mod2$failure_matrix[,1],
     100*simulation_full_results_mod2$failure_matrix[,2],
     xlab='True PMQ failure rate', ylab = 'Estimated PMQ failure rate')
lines(c(0,1),c(0,1))
```


# Simulating seasonal variation

Estimate the empirical seasonal reinfecton distribution from enrollment episodes. This assumes that the majority of enrollment episodes are reinfections which should be approximately true.
```{r}
load('../RData/TimingModel/Combined_Time_Event.RData')
Combined_Time_Data = filter(Combined_Time_Data, Censored == -1)
ind = Combined_Time_Data$WeekTime>52
Combined_Time_Data$WeekTime[ind] = Combined_Time_Data$WeekTime[ind]-52
hist(Combined_Time_Data$WeekTime, breaks = 0:52, 
     main = 'Week of enrollment',xlab='Week of the year')
# the vector we can use for empirical distribution sampling
seasonal_sampling_vector = Combined_Time_Data$WeekTime
```

Generate patient data under the assumptions of model 1 with the additional assumption of seasonality for reinfection.
```{r}
# test to show this does the right thing
set.seed(76576)
ys = sapply(1:1000, FUN = function(x, seasonal_sampling_vector){
  generate_reinfection_time_seasonal(params_M1,0,seasonal_sampling_vector)
}, seasonal_sampling_vector)
hist(ys, breaks = seq(0,10000,by=14), xlim = c(0,3*360), 
     main='Reinfection times',
     xlab='days from 1st January')
```


Generate data with these parameters:
```{r}
set.seed(87678)
out3 = simulate_dataset(N_PMQ = N_PMQ, N_CQ = N_CQ, 
                        N_AS = N_AS,FUP_time = FUP_time,
                        data_generation_function = generate_patient_data_Model2_Seasonal,
                        params = params_M2,
                        seasonal_sampling_vector = seasonal_sampling_vector)
Simdata_Model_Seasonal = out3$Simdata
Simulation_truth_Seasonal = out3$Simulation_truth
```


Fit stan model to simulated data where assumption of constant reinfection rate is wrong
```{r, include=FALSE}
if(RUN_MODELS){
  mod_seasonal_Fit = sampling(Timing_Model2,
                              data = c(Simdata_Model_Seasonal, Prior_params_M2),
                              iter = IT, warmup = WarmUp,
                              chains=Chains, thin = thin, 
                              init = lapply(1:Chains, 
                                            FUN = function(x,Prior_params_M2,N_noPMQ,N_PMQ){
                                              generate_inital_values_M2(Prior_params_M2,N_noPMQ,N_PMQ)
                                            }, 
                                            Prior_params = Prior_params_M2, 
                                            N_noPMQ = Simdata_Model_Seasonal$N_noPMQ,
                                            N_PMQ = Simdata_Model_Seasonal$N_PMQ))
  save(mod_seasonal_Fit, file = '../RData/LargeFiles/StanModels_SimSeasonal_mod2.RData')
} else {
  load(file = '../RData/LargeFiles/StanModels_SimSeasonal_mod2.RData')
}
thetas_mod_seasonal = extract(mod_seasonal_Fit)
```


Plot output, comparing prior, ground truth and model estimate:
```{r plotModel_Seasonal_fit}
plot_output_model1(thetas_mod_seasonal,Simulation_truth_Seasonal,
                   Simdata_Model_Seasonal,Prior_params_M1)
```

### Run multiple random iterations to assess model fit and bias


```{r, include=FALSE}
params_interest = c('lambda','gamma','logit_mean_p',
                    'logit_sd_p','logit_EarlyL')
Sim_Iterations = 100
if(RUN_MODELS){
  comparison_matrix = array(dim = c(Sim_Iterations,length(params_interest),2))
  for(sim in 1:Sim_Iterations){
    results = full_shebang(simulation_patient_function = 
                             generate_patient_data_Model2_Seasonal,
                           simulate_data_function = simulate_dataset,
                           stan_model = Timing_Model2,
                           params = params_M2,
                           init_values_function = generate_inital_values_M2,
                           N_AS = N_AS, N_CQ = N_CQ, N_PMQ = N_PMQ,
                           Prior_params = Prior_params_M2,
                           IT = IT, WarmUp = WarmUp, Chains = Chains, thin = thin,
                           params_interest = params_interest,
                           seasonal_sampling_vector=seasonal_sampling_vector)
    
    # record results: mean estimated value and true simulation value
    for(p in params_interest){
      comparison_matrix[sim,which(p==params_interest),] = results$summary_parameters[[p]]
    }
  }
  
  save(comparison_matrix,
       file = '../RData/TimingModel/simulation_full_results_mod_seasonal.RData')
} else {
  load('../RData/TimingModel/simulation_full_results_mod_seasonal.RData')
}
```



```{r full_shebang_results_mod2}
for(p in params_interest){
  writeLines(sprintf('For parameter %s, %s%% are below the true parameter',p,100*sum(comparison_matrix[,which(p==params_interest),1]<
                                                                                       unique(comparison_matrix[,which(p==params_interest),2]))/Sim_Iterations))
  hist(comparison_matrix[,which(p==params_interest),1], main=p)
  abline(v=unique(comparison_matrix[,which(p==params_interest),2]),col='red',lwd=3)
}

# plot(100*simulation_full_results_mod2$failure_matrix[,1],
#      100*simulation_full_results_mod2$failure_matrix[,2],
#      xlab='True PMQ failure rate', ylab = 'Estimated PMQ failure rate')
# lines(c(0,1),c(0,1))
```


# Posterior predictive p-values

In this section we simulate data from the posterior predictive for Model 2. We then apply a qualitative model check by comparing summary statistics from the posterior predicitive with the true data summary statistics.

```{r}
load('../RData/TimingModel/Posterior_theta_samples.RData')
load('../RData/TimingModel/Combined_Time_Event.RData')


PMQ_FUP = Combined_Time_Data$FU_time[!duplicated(Combined_Time_Data$patientid)&
                                       Combined_Time_Data$arm_num=='CHQ/PMQ']
CQ_FUP = Combined_Time_Data$FU_time[!duplicated(Combined_Time_Data$patientid)&
                                      Combined_Time_Data$arm_num=='CHQ']
AS_FUP = Combined_Time_Data$FU_time[!duplicated(Combined_Time_Data$patientid)&
                                      Combined_Time_Data$arm_num=='AS']
N_PMQ = sum(Combined_Time_Data$arm_num=='CHQ/PMQ'&
              !duplicated(Combined_Time_Data$patientid))
N_CQ = sum(Combined_Time_Data$arm_num=='CHQ'&
             !duplicated(Combined_Time_Data$patientid))
N_AS = sum(Combined_Time_Data$arm_num=='AS'&
             !duplicated(Combined_Time_Data$patientid))
N_PMQ_BPD = sum(Combined_Time_Data$arm_num=='CHQ/PMQ'&
                  Combined_Time_Data$Study_Period==2 &
                  !duplicated(Combined_Time_Data$patientid))

set.seed(45758)
params_M2_posterior = list()
i = sample(1:nrow(thetas2_matrix),1)
for(p in names(params_M2)){
  params_M2_posterior[[p]] = thetas2_matrix[i,p]
}

dat_posterior =
  simulate_dataset(N_PMQ = N_PMQ,N_CQ = N_CQ,N_AS = N_AS,
                   FUP_time = c(sample(x = PMQ_FUP,size = length(PMQ_FUP),replace = T),
                                sample(x = CQ_FUP,size = length(CQ_FUP),replace = T),
                                sample(x = AS_FUP,size = length(AS_FUP),replace = T)),
                   Study_Period = c(rep(2,N_PMQ_BPD),rep(2,N_PMQ-N_PMQ_BPD),rep(1,N_CQ+N_AS)),
                   data_generation_function = generate_patient_data_Model2,
                   params = params_M2_posterior)

AS_ind = dat_posterior$Simdata$Drug==0 & 
  dat_posterior$Simdata$Censored==0
AS_ind_true = Combined_Time_Data$arm_num=='AS'&
  Combined_Time_Data$Censored==0
CQ_ind = dat_posterior$Simdata$Drug==1 & 
  dat_posterior$Simdata$Censored==0
CQ_ind_true = Combined_Time_Data$arm_num=='CHQ'&
  Combined_Time_Data$Censored==0
PMQ_ind = dat_posterior$Simdata$Drug==2 & dat_posterior$Simdata$Censored==0
PMQ_ind_true = Combined_Time_Data$arm_num=='CHQ/PMQ'&
  Combined_Time_Data$Censored==0

par(las=1, mfrow=c(1,2))
#** Artesunate monotherapy: posterior predictive versus true data **
hist(Combined_Time_Data$Time_to_event[AS_ind_true], 
     main = 'AS monotherapy', 
     xlab='Days from last episode', col=alpha('blue',alpha = .4), 
     breaks = seq(0,700,by=7), xlim = c(0,300))
hist(dat_posterior$Simdata$Durations[AS_ind],
     add=T, breaks = seq(0,700,by=7), col=alpha('red',alpha = .4))

plot.ecdf(Combined_Time_Data$Time_to_event[AS_ind_true],
          col='blue',verticals = T, xlab='Days from last episode',
          main='Time to recurrence after AS monotherapy')
plot.ecdf(dat_posterior$Simdata$Durations[AS_ind],
          add=T,col='red',verticals = T)
legend('right',col=c('red','blue'),
       legend = c('Simulated data from posterior predictive',
                  'Observed trial data'),lwd=2, inset = 0.01)

#** Chloroquine monotherapy: posterior predictive versus true data **
hist(Combined_Time_Data$Time_to_event[CQ_ind_true], main = 'CQ monotherapy', 
     xlab='days from last episode', col=alpha('blue',alpha = .4), 
     breaks = seq(0,700,by=7), xlim = c(0,300))
hist(dat_posterior$Simdata$Durations[CQ_ind],
     add=T, breaks = seq(0,700,by=7), col=alpha('red',alpha = .4))

plot.ecdf(Combined_Time_Data$Time_to_event[CQ_ind_true],
          col='blue',verticals = T, xlab='Days from last episode',
          main='Time to recurrence after CQ monotherapy')
plot.ecdf(dat_posterior$Simdata$Durations[CQ_ind],
          add=T,col='red',verticals = T)
legend('right',col=c('red','blue'),
       legend = c('Simulated data from posterior predictive',
                  'Observed trial data'),lwd=2, inset = 0.01)

#** Primaquine+ : posterior predictive versus true data **
hist(Combined_Time_Data$Time_to_event[PMQ_ind_true], main = 'PMQ+', 
     xlab='days from last episode', col=alpha('blue',alpha = .4), 
     breaks = seq(0,700,by=7), xlim = c(0,300))
hist(dat_posterior$Simdata$Durations[PMQ_ind],
     add=T, breaks = seq(0,700,by=7), col=alpha('red',alpha = .4))

plot.ecdf(Combined_Time_Data$Time_to_event[PMQ_ind_true],
          col='blue',verticals = T, xlab='Days from last episode',
          main='Time to recurrence after PMQ+')
plot.ecdf(dat_posterior$Simdata$Durations[PMQ_ind],
          add=T,col='red',verticals = T)
legend('right',col=c('red','blue'),
       legend = c('Simulated data from posterior predictive',
                  'Observed trial data'),lwd=2, inset = 0.01)
```

```{r posterior_pred_pvalues}
Sim_Iterations = 200
res = array(dim = c(Sim_Iterations,3))
pb = txtProgressBar(min=1, max = Sim_Iterations,style = 3)
for(Sim in 1:Sim_Iterations){
  
  params_M2_posterior = list()
  i = sample(1:nrow(thetas2_matrix),1)
  for(p in names(params_M2)){
    params_M2_posterior[[p]] = thetas2_matrix[i,p]
  }
  dat_posterior =
    simulate_dataset(N_PMQ = N_PMQ,N_CQ = N_CQ,N_AS = N_AS,
                     FUP_time = c(sample(x = PMQ_FUP,size = length(PMQ_FUP),replace = T),
                                  sample(x = CQ_FUP,size = length(CQ_FUP),replace = T),
                                  sample(x = AS_FUP,size = length(AS_FUP),replace = T)),
                     Study_Period = c(rep(2,N_PMQ_BPD),rep(2,N_PMQ-N_PMQ_BPD),rep(1,N_CQ+N_AS)),
                     data_generation_function = generate_patient_data_Model2,
                     params = params_M2_posterior)
  
  AS_ind = dat_posterior$Simdata$Drug==0 & 
    dat_posterior$Simdata$Censored==0
  CQ_ind = dat_posterior$Simdata$Drug==1 & 
    dat_posterior$Simdata$Censored==0
  PMQ_ind = dat_posterior$Simdata$Drug==2 & dat_posterior$Simdata$Censored==0
  
  # compute recurrences per follow-up year
  res[Sim,1] = length(dat_posterior$Simdata$Durations[AS_ind])/
    (sum(dat_posterior$Simdata$Durations[dat_posterior$Simdata$Drug==0])/360)
  res[Sim,2] = length(dat_posterior$Simdata$Durations[CQ_ind])/
    (sum(dat_posterior$Simdata$Durations[dat_posterior$Simdata$Drug==1])/360)
  res[Sim,3] = length(dat_posterior$Simdata$Durations[PMQ_ind])/
    (sum(dat_posterior$Simdata$Durations[dat_posterior$Simdata$Drug==2])/360)
  setTxtProgressBar(pb,value = Sim)
}
```

```{r}
par(mfrow=c(1,3))
hist(res[,1], xlab = 'Recurrences per person (%)', main = 'Artesunate monotherapy')
true_stat_AS = length(Combined_Time_Data$Time_to_event[AS_ind_true])/
    (sum(Combined_Time_Data$Time_to_event[Combined_Time_Data$arm_num=='AS'])/360)
abline(v = true_stat_AS, col='blue',lwd=3)

hist(res[,2], xlab = 'Recurrences per person (%)', main = 'Chloroquine monotherapy')
true_stat_CQ = length(Combined_Time_Data$Time_to_event[CQ_ind_true])/
    (sum(Combined_Time_Data$Time_to_event[Combined_Time_Data$arm_num=='CHQ'])/360)
abline(v = true_stat_CQ,  col='blue',lwd=3)

hist(res[,3], xlab = 'Recurrences per person (%)', main = 'Primaquine+')
true_stat_PMQ = length(Combined_Time_Data$Time_to_event[PMQ_ind_true])/
    (sum(Combined_Time_Data$Time_to_event[Combined_Time_Data$arm_num=='CHQ/PMQ'])/360)
abline(v = true_stat_PMQ,col='blue',lwd=3)

PP_pval_AS = 1-sum(res[,1]>true_stat_AS)/Sim_Iterations
writeLines(sprintf('The posterior predictive p value for the number of recurrences per person in the AS arm is %s',PP_pval_AS))
PP_pval_CQ = 1-sum(res[,2]>true_stat_CQ)/Sim_Iterations
writeLines(sprintf('The posterior predictive p value for the number of recurrences per person in the CQ arm is %s',PP_pval_CQ))
PP_pval_PMQ = 1-sum(res[,3]>true_stat_PMQ)/Sim_Iterations
writeLines(sprintf('The posterior predictive p value for the number of recurrences per person in the PMQ+ arm is %s',PP_pval_PMQ))
```

