---
title: "Simulation Study for Timing Model"
author: "James Watson"
date: "07/05/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE, cache.comments = FALSE, 
                      echo = TRUE, include = TRUE, 
                      fig.width = 7, fig.height = 7,
                      fig.pos = 'H', 
                      dev = c('pdf','png'), dpi = 300)

load('../RData/RPackages_List.RData')
new.pkg <- pkgs[!(pkgs %in% installed.packages()[, "Package"])]
if(length(new.pkg) > 0){ # Install using .R script
  stop('Please run the script Install_and_load_required_packages.R before returning to this script. Thanks.')}else{
    sapply(pkgs, require, character.only = TRUE) # Load all packages
  }
```


```{r, echo=FALSE}
# Note that recompilation is necessary on different platforms
RECOMPILE_MODELS = F
# Running the full stan models can take several days
RUN_MODELS = T
```


## Simulated data

We explore a few scenarios of increasing complexity to check:

* The model can recover correct parameters when the data generating process is correctly specified (sanity check)
* The effect of a mis-specified data generating process. For this we look at the impact of seasonality

Functions to generate data
```{r}
generate_relapse_time = function(drug, params){
  early_versus_late = runif(1)
  type=NA
  if(early_versus_late < params$EarlyL){ # early relapse
    if(drug %in% c('PMQ/CHQ', 'CHQ')){
      next_event = round(rweibull(1, shape = params$CQ_shape, scale = params$CQ_scale))
    } else if (drug=='AS') {
      next_event = round(rweibull(1, shape = params$AS_shape, scale = params$AS_scale))
    }
    type = 'EarlyRelapse'
  } else {
    next_event = round(rexp(1, rate = params$gamma))
    type = 'LateRelapse'
  }
  return(list(time=next_event,type=type))
}

generate_reinfection_time = function(params){
  next_event = round(rexp(1, rate = params$lambda))
  return(next_event)
}

generate_recrudescence_time = function(drug, params){
  next_event = round(rexp(1, rate = params$lambda_recrud))
  return(next_event)
}

generate_patient_data_Model1 = function(params, follow_up=360, drug, ID=1){
  time_events = c()
  censor_status = c()
  true_event_type = c()
  total_time = 0
  passed_EOF = FALSE # passed end of follow-up
  # randomly generate the propensity to relapse
  p = inv.logit(rnorm(1,mean = params$logit_mean_p,sd = params$logit_sd_p))
  if(drug=='CHQ') { c1 = params$c1_CQ } else if(drug=='AS') { c1 = params$c1_AS}
  
  R_event = NA
  while(!passed_EOF){
    if(drug == 'CHQ/PMQ'){ 
      # primaquine: 100% efficacy, reinfection is the only option
      next_event = generate_reinfection_time(params = params)
      R_event = 'Reinfection'
    } else if (drug %in% c('CHQ','AS')){
      if(runif(1) < p){
        next_event = generate_reinfection_time(params)
        R_event = 'Reinfection'
      } else {
        if(runif(1) < c1){ 
          next_event = generate_recrudescence_time(drug = drug, params = params)
          R_event = 'Recrudescence'
        } else {
          lapse = generate_relapse_time(drug = drug, params = params)
          next_event = lapse$time
          R_event = lapse$type
        }
      }
    }
    
    total_time = total_time + next_event
    
    if(total_time <= follow_up){
      censor_status = c(censor_status, 0)
      time_events = c(time_events, next_event)
      true_event_type = c(true_event_type, R_event)
    } else {
      censor_status = c(censor_status, 1)
      passed_EOF = TRUE
      time_events = c(time_events,follow_up - (total_time-next_event))
      true_event_type = c(true_event_type, R_event)
    }
  }
  
  results = data.frame(time_events=time_events, censor_status=censor_status,
                       ID=ID, drug=drug, true_p = p,
                       true_event_type = true_event_type)
  return(results)
}
```


# Data simulation for model 1

Choose simulation parameters:
```{r}
params = list(lambda = 1/900,
              gamma = 1/50,
              lambda_recrud = 1/10,
              EarlyL = 0.5,
              c1_CQ = 0.01,
              c1_AS = 0.01,
              logit_mean_p = logit(0.2),
              logit_sd_p = (-logit(0.2)+logit(0.8))/1.96,
              AS_shape = 10,
              AS_scale = 21,
              CQ_shape = 10,
              CQ_scale = 42)
```


Generate a dataset
```{r}
#set.seed(475732)
drugs = c('CHQ/PMQ','CHQ','AS')
Ns = c(1000, 300, 300)
FUP_time = 300 # the follow-up time

N = sum(Ns)
id = 1
sim_data_mod1 = data.frame(time_events=NA, censor_status=NA,
                           ID=NA, drug=NA, true_p=NA,true_event_type=NA)
for(i in 1:length(drugs)){
  drug=drugs[i]
  if(Ns[i]>0){
    for(j in 1:Ns[i]){
      dat = generate_patient_data_Model1(params = params,
                                         follow_up = FUP_time,
                                         drug = drug,
                                         ID=id)
      id=id+1
      sim_data_mod1 = rbind(sim_data_mod1,dat)
    }
  }
  
}
sim_data_mod1 = sim_data_mod1[-1,]
sim_data_mod1 = filter(sim_data_mod1, !(censor_status > -1 & time_events < 5))


N_noPMQ = sum(sim_data_mod1$drug[!duplicated(sim_data_mod1$ID)] %in% c('CHQ','AS'))
N_PMQ = sum(sim_data_mod1$drug[!duplicated(sim_data_mod1$ID)] == 'CHQ/PMQ')
# Turn drug into a numeric vector
sim_data_mod1$numeric_drug = as.integer(revalue(sim_data_mod1$drug,
                                                c('AS'='0','CHQ'='1','CHQ/PMQ'='2')))


# Create a vector that maps the nth person to 0 (always received Primaquine) 
# or to their rank for those who didn't receive primaquine
noPMQ_ind = which(sim_data_mod1$drug != "CHQ/PMQ")
N_noPMQ = length(unique(sim_data_mod1$ID[noPMQ_ind])) #number of IDs without PMQ
ID_mapped_to_noPMQ_rank = rep(0, N)

index = 1
for(id in 1:N){
  ind = which(sim_data_mod1$ID==id)
  if(!2 %in% sim_data_mod1$numeric_drug[ind]){
    ID_mapped_to_noPMQ_rank[id] = index
    index = index + 1
  }
}

table(sim_data_mod1$true_event_type[sim_data_mod1$censor_status==0&sim_data_mod1$numeric_drug<2])
table(sim_data_mod1$true_event_type[sim_data_mod1$censor_status==1])


table(sim_data_mod1$true_event_type[sim_data_mod1$numeric_drug<2])
table(sim_data_mod1$true_event_type[sim_data_mod1$numeric_drug==2])
```

```{r}
par(las=1, mfrow=c(2,2))
reinfection_times = sapply(1:1000,function(x,params) generate_reinfection_time(params), 
                           params = params)
reinfection_times = reinfection_times[reinfection_times<FUP_time]

hist(reinfection_times,main = 'Times of reinfection')

recrudescence_times = sapply(1:1000,function(x,params) 
  generate_recrudescence_time(drug = 'CHQ',params=params), 
  params = params)
hist(recrudescence_times,main = 'Times of recrudescence')


early_relapsesAS = sapply(1:1000,
                          FUN = function(x,params) {
                            lapse = generate_relapse_time(params=params,drug = 'AS')
                            if(lapse$type=='EarlyRelapse') return(lapse$time) else return(NA)
                          },
                          params = params)
early_relapsesAS = early_relapsesAS[complete.cases(early_relapsesAS)]
early_relapsesCQ = sapply(1:1000,
                          FUN = function(x,params) {
                            lapse = generate_relapse_time(params=params,drug = 'CHQ')
                            if(lapse$type=='EarlyRelapse') return(lapse$time) else return(NA)
                          },
                          params = params)
early_relapsesCQ = early_relapsesCQ[complete.cases(early_relapsesCQ)]

plot(density(early_relapsesCQ), col='red',lwd=3, xlim=c(10,60))
lines(density(early_relapsesAS), col='green',lwd=3)

late_relapsesAS = sapply(1:1000,
                         FUN = function(x,params) {
                           lapse = generate_relapse_time(params=params,drug = 'AS')
                           if(lapse$type=='LateRelapse') return(lapse$time) else return(NA)
                         },
                         params = params)

late_relapsesCQ = sapply(1:1000,
                         FUN = function(x,params) {
                           lapse = generate_relapse_time(params=params,drug = 'CHQ')
                           if(lapse$type=='LateRelapse') return(lapse$time) else return(NA)
                         },
                         params = params)

hist(c(late_relapsesCQ, late_relapsesAS), xlim = c(0,FUP_time))
```


# Run Model1 on simulated data generated from Model 1



```{r}
if(RECOMPILE_MODELS){
  writeLines('Compiling model 1....')
  source('../Timing_Model/StanModel1.R')
  save(Timing_Model1, file = '../RData/TimingModel/Timing_Model1.RData')
} else {
  load('../RData/TimingModel/Timing_Model1.RData')
}
```

Prior specification
```{r}
# The hierachical parameters defining the prior distributions for model 1
Prior_params_M0 = list(Hyper_lambda_scale = 100,
                       Hyper_lambda_rate = 100*(1/params$lambda),
                       Hyper_lambda_recrud_scale = 100,
                       Hyper_lambda_recrud_rate = 100*(1/params$lambda_recrud),
                       Hyper_AS_shape_mean = params$AS_shape,
                       Hyper_AS_shape_sd = 1,
                       Hyper_AS_scale_mean = params$AS_scale,
                       Hyper_AS_scale_sd = 1,
                       Hyper_CQ_shape_mean = params$CQ_shape,
                       Hyper_CQ_shape_sd = 1,
                       Hyper_CQ_scale_mean = params$CQ_scale,
                       Hyper_CQ_scale_sd = 2,
                       Hyper_logit_mean_p = params$logit_mean_p,
                       Hyper_logit_sd_p = 1,
                       Hyper_logit_c1_mean = logit(params$c1_AS),
                       Hyper_logit_c1_sd = .25)
# Model 1 has the same parameters with a few extra
Prior_params_M1 = c(Prior_params_M0, 
                    Early_L_logit_mean = logit(params$EarlyL),
                    Early_L_logit_sd = 1,
                    Hyper_gamma_scale = 100,
                    Hyper_gamma_rate = 100*(1/params$gamma))
```

Setup options for stan and dataset in stan format
```{r}
Chains = 1
options(mc.cores = Chains)
IT = 10^4
WarmUp = .5*IT
thin = 30

# put the data into stan format
# For model 1 
Simdata_Model1 =list(N         = N,
                     #Number of individuals
                     Neps      = as.integer(nrow(sim_data_mod1)),
                     #Number of durations
                     N_noPMQ   = as.integer(N_noPMQ),
                     # Number of individuals who do not recieve PMQ
                     N_PMQ     = as.integer(N_PMQ),
                     # Number of individuals who do not recieve PMQ
                     Durations = as.double(sim_data_mod1$time_events),
                     # Time to recurrence or time to censoring
                     Censored  = as.integer(sim_data_mod1$censor_status),
                     # If the duration is right censored or not
                     Drug      = sim_data_mod1$numeric_drug,
                     # drug coded as an integer
                     ID_of_Patient = sim_data_mod1$ID,
                     # the ID corresponding to each time interval
                     ID_mapped_to_noPMQ_rank = ID_mapped_to_noPMQ_rank
                     # the index mapping PMQ individuals to their rank
)
```

Run model on simulated data
```{r, include=FALSE}
if(RUN_MODELS){
  mod1_Fit = sampling(Timing_Model1,
                      data = c(Simdata_Model1, Prior_params_M1),
                      iter = IT, warmup = WarmUp,
                      chains=Chains, thin = thin)
  save(mod1_Fit, file = '../RData/LargeFiles/StanModels_Sim1_mod1.RData')
}
thetas_mod1 = extract(mod1_Fit)
```

```{r plotModel1}
par(las=1)
par(mfrow=c(2,2))

# Time to reinfection
hist(1/thetas_mod1$lambda, main='Mean time to reinfection', xlab='1/lambda (days)')
abline(v=params$inv_lambda,col='red')
# Time to late relapse
hist(1/thetas_mod1$gamma, main='Mean time to late reLapse', xlab='1/gamma (days)')
abline(v=params$inv_gamma,col='red')
# Proportion early/late relapse
hist((thetas_mod1$logit_EarlyL), main='Logit early relapse', xlab='')
abline(v=logit(params$EarlyL),col='red')

par(mfrow=c(2,2))
# recrudescence proportion
hist(thetas_mod1$logit_c1_AS, main='Logit c1 AS', xlab='')
abline(v=logit(params$c1_AS),col='red')
hist(thetas_mod1$logit_c1_CQ, main='Logit c1 CQ', xlab='')
abline(v=logit(params$c1_CQ),col='red')
hist(thetas_mod1$Recrud_shape, main='Recrud shape', xlab='')
abline(v=(params$Recrud_shape),col='red')
hist(thetas_mod1$Recrud_scale, main='Recrud scale', xlab='')
abline(v=(params$Recrud_scale),col='red')

par(mfrow=c(2,2))
# shape and scale for AS early relapse
hist(thetas_mod1$AS_shape, main='AS shape', xlab='')
abline(v=(params$AS_shape),col='red')
hist(thetas_mod1$AS_scale, main='AS scale', xlab='')
abline(v=(params$AS_scale),col='red')
# shape and scale for CQ early relapse
hist(thetas_mod1$CQ_shape, main='CQ shape', xlab='')
abline(v=(params$CQ_shape),col='red')
hist(thetas_mod1$CQ_scale, main='CQ scale', xlab='')
abline(v=(params$CQ_scale),col='red')

par(mfrow=c(2,2))
# proportion reinfection
hist(inv.logit(thetas_mod1$logit_mean_p), main='Logit mean p', xlab='')
abline(v=inv.logit(params$logit_mean_p),col='red')
hist(thetas_mod1$logit_sd_p, main='Logit sd p', xlab='')
abline(v=(params$logit_sd_p),col='red')
# individual estimates versus true params for proportion reinfection
p_estimates = (apply(thetas_mod1$logit_p,2,mean))
true_p_estimates= logit(sim_data_mod1$true_p[!duplicated(sim_data_mod1$ID) & sim_data_mod1$numeric_drug<2])
plot(p_estimates, true_p_estimates)
lines(c(-10,10),c(-10,10))
plot(p_estimates, p_estimates-true_p_estimates)
abline(h=0)
```

# Not assuming 100% efficacy of primaquine

The following function generates data under the assumptions of model 2 - primaquine does not have 100% efficacy

```{r}
generate_patient_data_Model2 = function(params, follow_up=360, drug, ID=1){
  time_events = c()
  censor_status = c()
  true_event_type = c()
  total_time = 0
  passed_EOF = FALSE # passed end of follow-up
  # randomly generate the propensity to relapse
  p = inv.logit(rnorm(1,mean = params$logit_mean_p,sd = params$logit_sd_p))
  p_PMQ = inv.logit(rnorm(1,mean = params$logit_mean_p_PMQ,sd = params$logit_sd_p_PMQ))
  if(drug=='CHQ' | drug == 'CHQ/PMQ') { 
    c1 = params$c1_CQ 
  } else if(drug=='AS') { 
    c1 = params$c1_AS
  }
  
  while(!passed_EOF){
    if(drug == 'CHQ/PMQ'){ 
      true_p = p_PMQ
      if(runif(1) < p_PMQ){ #reinfection
        next_event = generate_reinfection_time(params)
        R_event = 'Reinfection'
      } else {
        if(runif(1) < c1){ # recrudescence
          next_event = generate_recrudescence_time(drug = drug, params = params)
          R_event = 'Recrudescence'
        } else { # relapse
          lapse = generate_relapse_time(drug = drug, params = params)
          next_event = lapse$time
          R_event = lapse$type
        }
      }
      
    } else if (drug %in% c('CHQ','AS')){
      true_p = p
      if(runif(1) < p){ #reinfection
        next_event = generate_reinfection_time(params)
        R_event = 'Reinfection'
      } else {
        if(runif(1) < c1){ #recrudescence
          next_event = generate_recrudescence_time(drug = drug, params = params)
          R_event = 'Recrudescence'
        } else { #relapse
          lapse = generate_relapse_time(drug = drug, params = params)
          next_event = lapse$time
          R_event = lapse$type
        }
      }
    }
    
    total_time = total_time + next_event
    
    if(total_time <= follow_up){
      censor_status = c(censor_status, 0)
      time_events = c(time_events, next_event)
      true_event_type = c(true_event_type, R_event)
    } else {
      censor_status = c(censor_status, 1)
      passed_EOF = TRUE
      time_events = c(time_events,follow_up - (total_time-next_event))
      true_event_type = c(true_event_type, NA)
    }
  }
  
  results = data.frame(time_events=time_events, censor_status=censor_status,
                       ID=ID, drug=drug, true_p = true_p,
                       true_event_type = true_event_type)
  return(results)
}
```

Generate some data
```{r}
params2 = list(inv_lambda = 900,
              inv_gamma = 50,
              EarlyL = 0.1,
              c1_CQ = 0.01,
              c1_AS = 0.01,
              logit_mean_p = logit(0.2),
              logit_sd_p = (-logit(0.2)+logit(0.8))/1.96,
              logit_mean_p_PMQ = logit(0.95),
              logit_sd_p_PMQ = (-logit(0.7)+logit(0.99))/1.96,
              Recrud_shape = 10,
              Recrud_scale = 15,
              AS_shape = 10,
              AS_scale = 21,
              CQ_shape = 10,
              CQ_scale = 42)

sim_data_mod2 = data.frame(time_events=NA, censor_status=NA,
                           ID=NA, drug=NA, true_p=NA,true_event_type=NA)
for(i in 1:length(drugs)){
  drug=drugs[i]
  if(Ns[i]>0){
    for(j in 1:Ns[i]){
      dat = generate_patient_data_Model2(params = params2,
                                         follow_up = FUP_time,
                                         drug = drug,
                                         ID=id)
      id=id+1
      sim_data_mod2 = rbind(sim_data_mod2,dat)
    }
  }
  
}
sim_data_mod2 = sim_data_mod2[-1,]
sim_data_mod2 = filter(sim_data_mod2, !(censor_status > -1 & time_events < 5))


N_noPMQ = sum(sim_data_mod2$drug[!duplicated(sim_data_mod2$ID)] %in% c('CHQ','AS'))
N_PMQ = sum(sim_data_mod2$drug[!duplicated(sim_data_mod2$ID)] == 'CHQ/PMQ')
# Turn drug into a numeric vector
sim_data_mod2$numeric_drug = as.integer(revalue(sim_data_mod2$drug,
                                                c('AS'='0','CHQ'='1','CHQ/PMQ'='2')))


# Create a vector that maps the nth person to 0 (always received Primaquine) 
# or to their rank for those who didn't receive primaquine
noPMQ_ind = which(sim_data_mod2$drug != "CHQ/PMQ")
N_noPMQ = length(unique(sim_data_mod2$ID[noPMQ_ind])) #number of IDs without PMQ
ID_mapped_to_noPMQ_rank = rep(0, N)

index = 1
for(id in 1:N){
  ind = which(sim_data_mod2$ID==id)
  if(!2 %in% sim_data_mod2$numeric_drug[ind]){
    ID_mapped_to_noPMQ_rank[id] = index
    index = index + 1
  }
}

table(sim_data_mod2$true_event_type[sim_data_mod2$numeric_drug<2])
table(sim_data_mod2$true_event_type[sim_data_mod2$numeric_drug==2])
```

# Simulating seasonal variation

```{r}
# include seasonal variation
generate_patient_data_Model3 = function(params, follow_up=360, drug, ID = 1){
  time_events = c()
  censor_status = c()
  total_time = 0
  passed_EOF = FALSE # passed end of follow-up
  # randomly generate the propensity to relapse
  p = inv.logit(rnorm(1,mean = params$logit_mean_p,sd = params$logit_sd_p))
  
  p_PMQ = inv.logit(rnorm(1,mean = params$logit_mean_p_PMQ,sd = params$logit_sd_p_PMQ))
  if(drug=='CHQ' | drug == 'CHQ/PMQ') { 
    c1 = params$c1_CQ 
  } else if(drug=='AS') { 
    c1 = params$c1_AS
  }
  
  while(!passed_EOF){
    if(drug == 'CHQ/PMQ'){ 
      if(runif(1) < p){
        next_event = generate_reinfection_time(params)
      } else {
        if(runif(1) < c1){ 
          next_event = generate_recrudescence_time(drug = drug, params = params)
        } else {
          next_event = generate_relapse_time(drug = drug, params = params)
        }
      }
      
    } else if (drug %in% c('CHQ','AS')){
      if(runif(1) < p){
        next_event = generate_reinfection_time(params)
      } else {
        if(runif(1) < c1){ 
          next_event = generate_recrudescence_time(drug = drug, params = params)
        } else {
          next_event = generate_relapse_time(drug = drug, params = params)
        }
      }
    }
    
    total_time = total_time + next_event
    
    if(total_time <= follow_up){
      censor_status = c(censor_status, 0)
      time_events = c(time_events, next_event)
    } else {
      censor_status = c(censor_status, 1)
      passed_EOF = TRUE
      time_events = c(time_events,follow_up - (total_time-next_event))
    }
  }
  
  results = data.frame(time_events=time_events, censor_status=censor_status,ID=ID, drug=drug)
  return(results)
}
```