---
title: "Pooled Analysis - supplementary"
author: "Aimee Taylor and James Watson"
output:
  pdf_document: default
  html_document:
    df_print: paged
    keep_md: yes
---

# Preamble

Load R packages, functions and data.

```{r, echo=FALSE, include=FALSE}
#==========================================================================
# Set up
#==========================================================================
knitr::opts_chunk$set(cache = TRUE, cache.comments = FALSE, 
                      include = TRUE, 
                      fig.width = 7, fig.height = 7,
                      fig.pos = 'H', 
                      dev = 'png', dpi = 300)


rm(list = ls())
load('../RData/RPackages_List.RData')
new.pkg <- pkgs[!(pkgs %in% installed.packages()[, "Package"])]
if(length(new.pkg) > 0){ # Install using .R script
  stop('Please run the script Install_and_load_required_packages.R before returning to this script. Thanks.')}else{
     sapply(pkgs, require, character.only = TRUE) # Load all packages
  }

tic()
load('../RData/TimingModel/Combined_Time_Event.RData')
load('../RData/FinalRecurrenceEstimates.RData')
source("../Genetic_Model/Data_functions.R")
source('../Genetic_Model/iGraph_functions.R')
source("../Genetic_Model/post_prob_CLI.R") 
source("../Genetic_Model/test_Rn_compatible.R") 
source("../Genetic_Model/Data_Inflation_Functions.R")

# The pooled MS data from BPD and VHX
load('../RData/GeneticModel/MS_data_PooledAnalysis.RData')

# Booleans that describe generation of Markdown file. 
# The large jobs should be done on the cluster

# runs the once model on those that can be directly computed
RUN_MODELS_SINGLE_SIMPLE = F           # can do on laptop (5 minutes)
# runs the model for 100 random draws from the Time model posterior on those that can be directly computed
RUN_MODELS_FULL_POSTERIOR_SIMPLE = F    # can do on laptop (2 hours)
# runs the model for 100 random draws from the Time model posterior on those that cannot be directly computed
RUN_MODELS_FULL_POSTERIOR_INFLATED = F  # cluster is better (10 hours)
# Does the false positive discovery estimation
RUN_MODELS_FALSE_POSITIVE = F           # only cluster (1 day with 42 cores)
# generates plots
CREATE_PLOTS = T

# Colour scheme
# Previous Set1 not colourblind friendly: display.brewer.all(colorblindFriendly = T)
Dark2 = brewer.pal(8, 'Dark2')

# This needs moving to time analysis: 
drug_cols3 = array(Dark2[c(4,6,1)], dim = 3, dimnames = list(c('AS','CHQ','CHQ/PMQ'))) 
drug_cols2 = array(Dark2[c(2,2,1)], dim = 3, dimnames = list(c('AS','CHQ','CHQ/PMQ')))
drug_cols_light2 = sapply(drug_cols2, adjustcolor, alpha.f = 0.5)

# Vector of states
states = c(relapse = 'L', reinfection = 'I', recrudescence = 'C')

# James, there are some odd bugs something to do with !duplicated() and the tibble...
# https://stackoverflow.com/questions/39041115/fixing-a-multiple-warning-unknown-column

## Threshold values for classification
Epsilon_upper = 0.9
Epsilon_lower = 0.1
transparent_pink_band = adjustcolor(Dark2[4], alpha.f = 0.2)
```

## Data processing

```{r}
# First we remove MS data for which there are no recurrent data
N_episodes_typed = table(MS_pooled$ID[!duplicated(MS_pooled$Episode_Identifier)])
MS_pooled = filter(MS_pooled, ID %in% names(N_episodes_typed[N_episodes_typed>1]))
# recreate pooled summary dataset
MS_pooled_summary = MS_pooled[!duplicated(MS_pooled$Episode_Identifier),] 

writeLines(sprintf('Number of individuals with at least two typed episodes analysed: %s',
                   length(unique(MS_pooled$ID))))
writeLines(sprintf('Number of episodes in individuals with at least two typed episodes analysed: %s',
                   length(unique(MS_pooled$Episode_Identifier))))
writeLines(sprintf('Number of typed recurrences analysed: %s',
                   length(unique(MS_pooled$Episode_Identifier[MS_pooled$Episode>1]))))
```

First we define the set of microsatellite markers used in this analysis:
```{r}
MSs_all = c("PV.3.502","PV.3.27","PV.ms8",
            "PV.1.501","PV.ms1","PV.ms5",
            "PV.ms6","PV.ms7","PV.ms16")
```

We use a multinomial-dirichlet model with subjective weight $\omega$. $\omega = 0$ recovers the empirical allele frequencies. 

```{r, echo=F}
# Prior weight for the Dirichlet (setting weight to 0 recovers empirical freq):
D_weight_Prior = 1

# These are motif lengths: for the plotting
MSs_Motifs = list("PV.3.502"=8,'PV.3.27' = 4, 
                  "PV.ms8" = 3, "PV.1.501"= 7, 
                  "PV.ms1" = 3, "PV.ms5" = 3,
                  "PV.ms6" = 3, "PV.ms16" =3,
                  "PV.ms7" = 3)

writeLines(paste('Number of episodes used to compute frequencies:',
                 sum(MS_pooled$Episode==1 & MS_pooled$MOI_id==1)))
Ind_Primary = which(MS_pooled$Episode==1)

# Posterior Dirichlet parameter vector
Alpha_Posteriors = apply(MS_pooled[,MSs_all], 2, function(x, Ind_Primary){
  # Extract xmax 
  xmax = max(x,na.rm=T)
  # prior parameter vector (iterpolates unobserved repeat lengths < xmax)
  param_vector = array(D_weight_Prior, dim = xmax, dimnames = list(1:xmax)) 
  # observed data summarised as counts
  obs_counts = table(x[Ind_Primary]) 
  # posterior parameter vector
  param_vector[names(obs_counts)] = param_vector[names(obs_counts)] + obs_counts
  return(param_vector)
})

# Calculate posterior mean 
Fs_Combined = sapply(Alpha_Posteriors, function(x){x/sum(x)})
```


## Run model on easy individuals

```{r, include=FALSE}
if(RUN_MODELS_FULL_POSTERIOR_SIMPLE){
  #===============================================
  # Run full Bayesian with sampling at random from time prior 
  # (treats time posterior as discrete uniform prior)
  # takes about 1.1 hour on 6 cores
  # The outer loop is not parallelised - suboptimal coding
  #===============================================
  
  Ksamples = 100
  tic()
  Thetas_full_post_TAgnostic = foreach(ss = 1:Ksamples, .combine = cbind) %do% {
    
    # draw a random distribution over the allele frequencies from posterior
    Fs_random = lapply(Alpha_Posteriors, rdirichlet, n=1)
    # I'm not sure if this is needed - how are these called deep inside?
    for(i in 1:length(Fs_random)){
      names(Fs_random[[i]]) = 1:length(Fs_random[[i]])
    }
    
    # run the model
    thetas_9MS = post_prob_CLI(MSdata = MS_pooled, Fs = Fs_random, 
                               cores = 7, verbose = F) 
    thetas_9MS$Episode_Identifier = rownames(thetas_9MS)
    thetas_9MS
  }
  
  save(Thetas_full_post_TAgnostic, file = '../RData/GeneticModel/Full_Posterior_Model_samples_TAgnostic.RData')
  toc()
} else {
  load(file = '../RData/GeneticModel/Full_Posterior_Model_samples_TAgnostic.RData')
}
```


# Extra computations for VHX: complex individuals


We remove the IDs that can be straightforwardly calculated:

```{r}
ind_calculated = which(MS_pooled_summary$Episode_Identifier %in% Thetas_full_post_TAgnostic$Episode_Identifier)
IDs_calculated = unique(MS_pooled_summary$ID[ind_calculated])
IDs_remaining = unique(MS_pooled_summary$ID[! MS_pooled_summary$ID %in% IDs_calculated])
writeLines(sprintf('individuals with more than two recurrences: %s',length(IDs_remaining)))
```

We blow up the pooled analysis into all pairs within individuals:

```{r, include=FALSE}
MS_inflate = reformat_MSdata(filter(MS_pooled, ID %in% IDs_remaining), MSs = MSs_all)
MS_inflated = Inflate_into_pairs(MS_data = MS_inflate)
writeLines(sprintf('Number of pairwise probabilities of recurrence states between episodes: %s',
                   length(unique(MS_inflated$ID))))
```


```{r, include=FALSE}
lowerCI=0.05
upperCI=0.95
if(RUN_MODELS_FULL_POSTERIOR_INFLATED){  
  
  all_rec_eps = unique(MS_inflated$Episode_Identifier[MS_inflated$Episode==2])
  
  Ksamples = 100 # 11 hours with 6 cores
  
  K_results = sum(!duplicated(MS_inflated$Episode_Identifier[MS_inflated$Episode>1]))
  
  # Set up the prior matrix: 
  # recurrences have 0 prior probability if not comparing sequential episodes
  P_matrix = array(1/3,dim = c(nrow(MS_inflated),4) )
  colnames(P_matrix) = c("Episode_Identifier","C","L","I")
  P_matrix = data.frame(P_matrix)
  P_matrix[,'Episode_Identifier'] = MS_inflated$Episode_Identifier
  unique_recurrences = !duplicated(MS_inflated$Episode_Identifier) & MS_inflated$Episode>1
  ind_one_half = unique_recurrences & 
    MS_inflated$First_EpNumber < (MS_inflated$Second_EpNumber-1)
  P_matrix$C[ind_one_half] = 0
  P_matrix$L[ind_one_half] = 0.5
  P_matrix$I[ind_one_half] = 0.5
  P_matrix = P_matrix[unique_recurrences, ]
  
  tic()
  Res_total_TAgnostic=foreach(ss = 1:Ksamples, .combine = cbind) %do% {
    
    
    # draw a random distribution over the allele frequencies from posterior
    Fs_random = lapply(Alpha_Posteriors, rdirichlet, n=1)
    # I'm not sure if this is needed - how are these called deep inside?
    for(i in 1:length(Fs_random)){
      names(Fs_random[[i]]) = 1:length(Fs_random[[i]])
    }
    
    # This is about 6-7 minutes per run
    Res_TAgnostic = post_prob_CLI(MSdata = MS_inflated, 
                                  Fs = Fs_random, p = P_matrix,
                                  UpperComplexity = 10^6, 
                                  verbose = F,
                                  cores = 7)
    Res_TAgnostic
  }
  save(Res_total_TAgnostic, 
       file = '../RData/LargeFiles/FullPosterior_INF_TAgnostic.bigRData') # save for safety
  toc()
  # Summarise the results, all CIs are 90%
  # Save this summary
  C_cols = grep('C',colnames(Res_total_TAgnostic))
  L_cols = grep('L',colnames(Res_total_TAgnostic))
  I_cols = grep('I',colnames(Res_total_TAgnostic))
  Results_Inflated_TAgnostic = 
    data.frame(Episode_Identifier = rownames(Res_total_TAgnostic),
               C_median = apply(Res_total_TAgnostic[,C_cols],1,
                                quantile,probs = 0.5,na.rm=T),
               C_min = apply(Res_total_TAgnostic[,C_cols],1,
                             quantile,probs = lowerCI,na.rm=T),
               C_max = apply(Res_total_TAgnostic[,C_cols],1,
                             quantile,probs = upperCI,na.rm=T),
               L_median = apply(Res_total_TAgnostic[,L_cols],1,
                                quantile,probs = 0.5,na.rm=T),
               L_min = apply(Res_total_TAgnostic[,L_cols],1,
                             quantile,probs = lowerCI,na.rm=T),
               L_max = apply(Res_total_TAgnostic[,L_cols],1,
                             quantile,probs = upperCI,na.rm=T),
               I_median = apply(Res_total_TAgnostic[,I_cols],1,
                                quantile,probs = 0.5,na.rm=T),
               I_min = apply(Res_total_TAgnostic[,I_cols],1,
                             quantile,probs = lowerCI,na.rm=T),
               I_max = apply(Res_total_TAgnostic[,I_cols],1,
                             quantile,probs = upperCI,na.rm=T))
  save(Results_Inflated_TAgnostic, 
       file = '../RData/GeneticModel/Full_Posterior_Inflated_TAgnostic.RData')
} else {
  load('../RData/GeneticModel/Full_Posterior_Inflated_TAgnostic.RData')
}
```


Construct adjacency graphs and compute probabilities of relapse and reinfection.

```{r}
MS_pooled_summary$L_or_C_state = MS_pooled_summary$TotalEpisodes = NA
MS_pooled_summary$L_lower = MS_pooled_summary$L_upper = MS_pooled_summary$L_median = NA
MS_pooled_summary$C_lower = MS_pooled_summary$C_upper = MS_pooled_summary$C_median = NA
MS_pooled_summary$I_lower = MS_pooled_summary$I_upper = MS_pooled_summary$I_median = NA
# Arrange by complexity
# Get single rows per episode (throw away the extra MOI information)
MS_inflated_summary = MS_inflated[!duplicated(MS_inflated$Episode_Identifier) & 
                                    MS_inflated$Episode==2,]
Results_Inflated_TAgnostic$Episode_Identifier = as.character(Results_Inflated_TAgnostic$Episode_Identifier)
for(i in 1:nrow(MS_inflated_summary)){
  if(!length(which(MS_inflated_summary$Episode_Identifier[i] == 
                   Results_Inflated_TAgnostic$Episode_Identifier))>0){
    MS_inflated_summary = MS_inflated_summary[-i,]
    print('removing')
  }
}

Results_Inflated_TAgnostic$ID_True = NA
Results_Inflated_TAgnostic$First_EpNumber = NA
Results_Inflated_TAgnostic$Second_EpNumber = NA
# The ordering has changed so need to be careful about naming
for(i in 1:nrow(Results_Inflated_TAgnostic)){
  ind_MS_inflated = which(MS_inflated_summary$Episode_Identifier==Results_Inflated_TAgnostic$Episode_Identifier[i])
  Results_Inflated_TAgnostic$ID_True[i] = 
    MS_inflated_summary$ID_True[ind_MS_inflated]
  Results_Inflated_TAgnostic$First_EpNumber[i] =
    MS_inflated_summary$First_EpNumber[ind_MS_inflated]
  Results_Inflated_TAgnostic$Second_EpNumber[i] =
    MS_inflated_summary$Second_EpNumber[ind_MS_inflated]
}



# Iterate through the ones we can calculate in one go
episodes_full_model = unique(Thetas_full_post_TAgnostic$Episode_Identifier)
cols_remove = grep('Episode_Identifier', colnames(Thetas_full_post_TAgnostic))
Thetas_full_post_TAgnostic = Thetas_full_post_TAgnostic[, -cols_remove]
for(ep in episodes_full_model){
  ind1 = (MS_pooled_summary$Episode_Identifier==ep)
  ind2 = rownames(Thetas_full_post_TAgnostic)==ep
  
  ## Summaries for relapse
  L_cols = grep('L',colnames(Thetas_full_post_TAgnostic))
  MS_pooled_summary$L_upper[ind1] = quantile(unlist(Thetas_full_post_TAgnostic[ind2,L_cols]),
                                             probs=upperCI, na.rm = T)
  MS_pooled_summary$L_lower[ind1] = quantile(unlist(Thetas_full_post_TAgnostic[ind2,L_cols]),
                                             probs=lowerCI, na.rm = T)
  MS_pooled_summary$L_median[ind1] = quantile(unlist(Thetas_full_post_TAgnostic[ind2,L_cols]),
                                              probs=0.5, na.rm = T)
  
  ## Summaries for recrudescence
  C_cols = grep('C',colnames(Thetas_full_post_TAgnostic))
  MS_pooled_summary$C_upper[ind1] = quantile(unlist(Thetas_full_post_TAgnostic[ind2,C_cols]),
                                             probs=upperCI, na.rm = T)
  MS_pooled_summary$C_lower[ind1] = quantile(unlist(Thetas_full_post_TAgnostic[ind2,C_cols]),
                                             probs=lowerCI, na.rm = T)
  MS_pooled_summary$C_median[ind1] = quantile(unlist(Thetas_full_post_TAgnostic[ind2,C_cols]),
                                              probs=0.5, na.rm = T)
  
  ## Summaries for reinfection
  I_cols = grep('I',colnames(Thetas_full_post_TAgnostic))
  MS_pooled_summary$I_upper[ind1] = quantile(unlist(Thetas_full_post_TAgnostic[ind2,I_cols]),
                                             probs=upperCI, na.rm = T)
  MS_pooled_summary$I_lower[ind1] = quantile(unlist(Thetas_full_post_TAgnostic[ind2,I_cols]),
                                             probs=lowerCI, na.rm = T)
  MS_pooled_summary$I_median[ind1] = quantile(unlist(Thetas_full_post_TAgnostic[ind2,I_cols]),
                                              probs=0.5, na.rm = T)
  
  # Just going to classify on relapse versus reinfection
  if(!is.na(MS_pooled_summary$L_upper[ind1])){
    if(MS_pooled_summary$L_upper[ind1]+MS_pooled_summary$C_upper[ind1] < Epsilon_lower){
      MS_pooled_summary$L_or_C_state[ind1] = 'I'
    } else if(MS_pooled_summary$L_lower[ind1]+MS_pooled_summary$C_lower[ind1] > Epsilon_upper){
      MS_pooled_summary$L_or_C_state[ind1] = 'L'
    } else {
      MS_pooled_summary$L_or_C_state[ind1] = 'Uncertain'
    }
  } else {
    MS_pooled_summary$L_or_C_state[ind1] = NA
  }
}
####******** Complex cases *******####################
# Now iterate through the complex ones
for(i in 1:length(IDs_remaining)){
  id = IDs_remaining[i]
  Doubles_Thetas = filter(Results_Inflated_TAgnostic, ID_True==id)
  
  for(ep in unique(Doubles_Thetas$Second_EpNumber)){
    # indices on the MS_pooled_summary
    ind1 = which(MS_pooled_summary$ID==id & MS_pooled_summary$Episode==ep)
    # indices on DOubles thetas: looking for relapse evidence
    ind2 = which(Doubles_Thetas$Second_EpNumber == ep)
    # index for recrudescence evidence
    ind3 = which(Doubles_Thetas$Second_EpNumber == ep &
                   Doubles_Thetas$First_EpNumber == (ep-1))
    
    best_match_relapse = which.max(Doubles_Thetas$L_median[ind2])
    if(length(best_match_relapse)>0){
      MS_pooled_summary$L_lower[ind1] = Doubles_Thetas$L_min[ind2[best_match_relapse]]
      MS_pooled_summary$L_upper[ind1] = Doubles_Thetas$L_max[ind2[best_match_relapse]]
      MS_pooled_summary$L_median[ind1] = Doubles_Thetas$L_median[ind2[best_match_relapse]]
      
      if(length(ind3)>0){
        MS_pooled_summary$C_lower[ind1] = Doubles_Thetas$C_min[ind3]
        MS_pooled_summary$C_upper[ind1] = Doubles_Thetas$C_max[ind3]
        MS_pooled_summary$C_median[ind1] = Doubles_Thetas$C_median[ind3]
      }
      MS_pooled_summary$I_lower[ind1] = Doubles_Thetas$I_min[ind2[best_match_relapse]]
      MS_pooled_summary$I_upper[ind1] = Doubles_Thetas$I_max[ind2[best_match_relapse]]
      MS_pooled_summary$I_median[ind1] = Doubles_Thetas$I_median[ind2[best_match_relapse]]
    }
    if(!is.na(MS_pooled_summary$C_median[ind1])){
      if(MS_pooled_summary$L_upper[ind1] < MS_pooled_summary$L_lower[ind1]){
        writeLines(sprintf('Problem with ID %s',id))
        stop()
      }
      if(MS_pooled_summary$L_upper[ind1]+MS_pooled_summary$C_upper[ind1] < Epsilon_lower){
        MS_pooled_summary$L_or_C_state[ind1] = 'I'
      } else if(MS_pooled_summary$L_lower[ind1]+MS_pooled_summary$C_lower[ind1] > Epsilon_upper){
        MS_pooled_summary$L_or_C_state[ind1] = 'L'
      } else {
        MS_pooled_summary$L_or_C_state[ind1] = 'Uncertain'
      }
    }
  }
}

MS_pooled_summary$Drug = MS_pooled_summary$FU = NA
for(id in MS_pooled_summary$ID){
  ind = MS_pooled_summary$ID==id
  MS_pooled_summary$TotalEpisodes[ind] = max(MS_pooled_summary$Episode[ind])
  MS_pooled_summary$Drug[ind] = as.numeric(
    Combined_Time_Data$arm_num[Combined_Time_Data$patientid==id][1] == 'CHQ/PMQ') + 2
  MS_pooled_summary$FU[ind] = Combined_Time_Data$FU_time[Combined_Time_Data$patientid==id][1]
}
MS_pooled_summary$Plotting_pch_Values = 
  as.numeric(mapvalues(MS_pooled_summary$L_or_C_state, from = c('L','Uncertain','I'), to = c(17,15,1)))
MS_pooled_summary$Plotting_col_Values = 
  as.numeric(mapvalues(MS_pooled_summary$L_or_C_state, from = c('L','Uncertain','I'), to = 1:3))
```


Coatney plot (genetic data informed only)

```{r CoatneyStylePLot, echo=FALSE}
# Only the recurrences for which we can compute estimates of recurrence state
MS_final = filter(MS_pooled_summary, !is.na(L_median))
MS_final = arrange(MS_final, desc(Drug), desc(FU), desc(TotalEpisodes))
ids = unique(MS_final$ID)

if(CREATE_PLOTS){
  ## Time series data colored by genetic STATE: classification
  mycols_states_bg = c('#FFFFFF',Dark2[4],'#000000') # colors for states - need uncertain ones as well
  mycols_states_fg = c(Dark2[8],Dark2[4],'#000000') # colors for states - need uncertain ones as well
  mycols_drugs = brewer.pal(n=3, name = 'Set1')
  par(mfrow = c(1,1), las=1, bty='n', cex.axis=.3, mar=c(4,0,1,1), family = 'serif')
  plot(NULL, ylab = '', xlab = '',
       xlim = c(0,370), 
       ylim = c(1,length(ids)),
       xaxt='n', yaxt='n')
  mtext(text = 'Months from start of study', side = 1, line=2.5, cex=1)
  axis(1, at = seq(0,370, by=60), labels = seq(0,370, by=60)/30, cex.axis=1)
  
  for(i in 1:length(ids)){
    
    id = ids[i]
    ind = which(MS_final$ID==id)
    
    # Add the follow up time line
    lines(c(0,MS_final$FU[ind[1]]), 
          c(i,i), lty=1, 
          lwd=.5, col=drug_cols2[MS_final$Drug[ind[1]]])
    
    # For highlighting long-latency
    # Add the follow up time line
    if(id == 'VHX_235' | id == 'BPD_27'){
      lines(c(0,MS_final$FU[ind[1]]), 
            c(i,i), lty = 'dashed', 
            lwd=.5, col= 'black')
      
      cols = mycols_states_fg[MS_final$Plotting_col_Values[ind]]
      points(MS_final$timeSinceEnrolment[ind], 
             rep(i,length(ind)), 
             pch=1,cex = 1.5, 
             col='black')
    }
    # Add points after highlight lines
    points(MS_final$timeSinceEnrolment[ind], rep(i,length(ind)), 
           pch=MS_final$Plotting_pch_Values[ind], 
           bg=mycols_states_bg[MS_final$Plotting_col_Values[ind]],
           col=mycols_states_fg[MS_final$Plotting_col_Values[ind]], 
           cex=.6)
    
    
    
  }
  lines(x = c(0,0), y = c(0,length(ids)),lwd=3)
  
  legend('topright', col = c(mycols_states_fg[c(3,2,1)]), 
         pch = c(17,15,1), cex = 1, bty = 'n', y.intersp = 0.75, title = 'Classification',
         legend = c('Relapse or Recrudescence','Uncertain','Reinfection'))
  
  legend('top', col = drug_cols2[2:3], 
         cex = 1, bty = 'n', y.intersp = 0.75, 
         lty = 1, lwd = 1,
         legend = c('No PMQ', 'PMQ+'))
}
```
```{r}
writeLines(sprintf('The Coatney style plot is showing %s recurrences in %s individuals',
                   nrow(MS_final), length(ids)))
```


Break down the results depending on whether PMQ was received or not:
```{r, echo=FALSE}
writeLines(sprintf('There are %s recurrences after PMQ+. The breakdown as %% of classification is as follows:', sum(MS_final$Treatment=='PMQ')))
round(100*table(MS_final$L_or_C_state[MS_final$Treatment=='PMQ'])/sum(MS_final$Treatment=='PMQ'))

writeLines(sprintf('\nThere are %s recurrences after no PMQ. The breakdown as %% of classification is as follows:', sum(MS_final$Treatment!='PMQ')))
round(100*table(MS_final$L_or_C_state[MS_final$Treatment!='PMQ'])/sum(MS_final$Treatment!='PMQ'))
toc()
```


